\chapter{Simple Groups}
Simple groups can be thought of as the `building blocks' of all (finite) groups. The finite simple groups have been completely classified; each belongs to one of 18 infinite families, or is one of 26 sporadic groups that do not follow a specific pattern. We look at the classification of the families of these simple groups here.

\section{Cyclic Groups of Prime Order}
The first infinite family of simple groups we will look at is the family of Cyclic Groups of Prime Order\index{cyclic group!of prime order}.

\begin{lemma}\label{lemma-cyclic-group-simple-iff-order-is-prime}
    $\Cn{n}$ is simple if and only if $n$ is prime.
\end{lemma}
\begin{proof}
    We first prove the forward direction. Suppose $\Cn{n}$ is simple with generator $g$. Then the only normal subgroups of $\Cn{n}$ are the trivial group and the group itself. Seeking a contradiction, assume $n$ is not prime; write $n = ab$ where $a$ and $b$ are positive integers that are both smaller than $n$. Then clearly $\langle g^a\rangle$ is a proper subgroup of $\Cn{n}$. Now $\Cn{n}$ is abelian (\myref{prop-cyclic-group-is-abelian}) which means all subgroups are normal (\myref{prop-subgroup-of-abelian-group-is-normal}). Hence we have found a non-trivial proper normal subgroup of $\Cn{n}$, namely $\langle g^a \rangle$, contradicting that $\Cn{n}$ has no non-trivial proper normal subgroups. Therefore $n$ is prime.

    We now prove the reverse direction. Suppose $n$ is a prime. Then by a corollary of Lagrange's theorem (\myref{corollary-group-with-prime-order-subgroups}), $\Cn{n}$ has no non-trivial proper subgroups. So the only subgroup with order smaller than $n$ is the trivial group, $\{e\}$. Clearly $\Cn{n}$ is normal in itself, and the trivial group is always a normal subgroup. Hence, as the only normal subgroups of $\Cn{n}$ are the trivial group and itself, thus $\Cn{n}$ is simple. Therefore, if $n$ is prime then $\Cn{n}$ is simple.
\end{proof}

In fact, we have a much stronger result which we prove here.

\begin{theorem}\label{thrm-abelian-group-simple-iff-cylic-group-of-prime-order}
    An abelian group is simple if and only if it has prime order.
\end{theorem}
Note that we do not assume that the abelian group is finite; we will show that the group is finite in the proof below.
\begin{proof}
    The reverse direction follows immediately from \myref{lemma-cyclic-group-simple-iff-order-is-prime}, so we prove the forward direction only.

    Suppose $G$ is a simple abelian group; we show that $G$ is finite. Let $g$ a non-identity element of $G$. Then $H = \langle g \rangle$ is a subgroup of $G$. In fact, since $G$ is abelian, $H \unlhd G$ (\myref{prop-subgroup-of-abelian-group-is-normal}). As $G$ is simple, therefore $H = G$, meaning that $g$ is a generator of $G$. Now if $G$ is an infinite group, then one also sees that $\langle g^2 \rangle < G$ which implies $\langle g^2 \rangle \lhd G$, contradicting the fact that $G$ is simple. Hence $G$ is a finite abelian group with generator $g$, meaning $G$ is cyclic. Result follows directly from \myref{lemma-cyclic-group-simple-iff-order-is-prime}.
\end{proof}

From this, we conclude that the only family of simple abelian groups is the family of cyclic groups of prime order.

\section{Alternating Group With Degree $>4$}
The other family of simple groups that is relatively easy to find (and define) is the family of alternating groups with degree above 4\index{alternating group!of degree $>4$}. However, to prove this claim, we need several preliminary results.

\begin{theorem}\label{thrm-group-of-order-60-with->1-sylow-5-subgroup-is-simple}
    Let $G$ be a group of order 60. If $G$ has more than one Sylow 5-subgroup then $G$ is simple.
\end{theorem}
\begin{proof}[Proof (see {\cite[Proposition 4.21]{dummit_foote_2004}})]
    By way of contradiction assume $G$ is a group of order 60 with more than one Sylow 5-subgroup, but has a non-trivial proper normal subgroup $H$. Note $60 = 5 \times 12$, so by the Third Sylow Theorem (\myref{thrm-sylow-3}),
    \begin{itemize}
        \item $12 \vert n_5$, so $n_5 \in \{1, 2, 3, 4, 6, 12\}$; and
        \item $n_5 \equiv 1 \pmod 5$, so $n_5 \in \{1, 6, 11, 16, \dots\}$.
    \end{itemize}
    Therefore $n_5 = 6$ as $n_5 > 1$ (given), i.e. there are 6 Sylow 5-subgroups.

    We note by Lagrange's theorem (\myref{thrm-lagrange}) that the order of $H$ belongs in the set $\{1, 2, 3, 4, 5, 6, 10, 12, 15, 20, 30, 60\}$. As $H$ is a non-trivial proper subgroup of $G$, thus $|H| \neq 1$ and $|H| \neq 60$. That leaves 4 cases which we will deal with separately.
    \begin{enumerate}
        \item $|H| = 6$. Note $6 = 2 \times 3$, so \myref{problem-group-of-order-pq-has-normal-subgroup-of-order-q} tells us that there exists a $N \lhd H$ with $|N| = 3$. Note also $[G:H] = 10$ which is not a multiple of 3, so \myref{problem-normal-subgroup-of-G-contains-all-sylow-p-subgroups} tells us that all Sylow 3-subgroups of $G$ are in $H$. But $N \lhd H$ means that $N$ is the unique Sylow 3-subgroup of $H$ and $G$ (\myref{corollary-sylow-subgroup-is-normal-if-it-is-unique}), so $N \lhd G$ (by the same corollary). Proceed to case 3, using $N$ in place of $H$.

        \item $|H| = 12$. Note $12 = 2^2 \times 3$. Now \myref{exercise-group-of-order-12-has-normal-subgroup-of-3-or-4} (later) tells us that there exists a normal subgroup of $H$ with order 3 or 4. Call that subgroup $N$. If $|N| = 3$ then it is a Sylow 3-subgroup; if $|N| = 4 = 2^2$ it is a Sylow 2-subgroup. As $N \lhd H$, thus $N$ is the unique Sylow 2- or 3- subgroup (\myref{corollary-sylow-subgroup-is-normal-if-it-is-unique}). Since $H \lhd G$, thus $H$ contains all Sylow 2- and 3-subgroups of $G$ (\myref{problem-normal-subgroup-of-G-contains-all-sylow-p-subgroups}), meaning $G$ has only one Sylow 2-subgroup or one Sylow 3-subgroup (or both), in particular $N$. Hence, $N \lhd G$ since a Sylow $p$-subgroup is unique if and only if it is normal (\myref{corollary-sylow-subgroup-is-normal-if-it-is-unique}). Proceed with case 3, using $N$ instead of $H$.

        \item $|H| \in \{2, 3, 4\}$. Since $H \lhd G$, thus $G/H$ is a group. Note $|G/H| \in \{15, 20, 30\}$. We claim that each of these cases produces a new normal subgroup of $G/H$ (call it $\bar{P}$) with order 5. This is proven for the case where $|G/H| = 30$ in \myref{problem-group-of-order-30-has-normal-subgroup-of-order-5}; the other two cases are for \myref{exercise-group-of-order-15-or-20-has-normal-subgroup-of-order-5} (later).

        Now \myref{problem-subgroup-of-quotient-group-is-quotient-group} tells us that $\bar{P}$ has the form $K/H$ where $K < G$ and $H \subseteq K$. Since $\bar{P} = K/H \lhd G/H$, thus for any $g \in G$ and $kH \in \bar{P}$ we have
        \[
            (gH)(kH)(g^{-1}H) = (gkg^{-1})H \in K/H,
        \]
        which means $gkg^{-1} \in K$. Therefore $K \lhd G$ by definition of normality.

        Observe that this means that
        \[
            |K| = |K/H||H| = |\bar{P}||H| = 5|H|,
        \]
        meaning $K$ is a normal subgroup of $G$ with an order that is a multiple of 5. Proceed to case 4, using $K$ in place of $H$.

        \item $|H|$ is a multiple of 5, meaning $H$ has a Sylow 5-subgroup. Note that there are $5-1=4$ non-identity elements in each Sylow 5-subgroup; therefore
        \[
            |H| \geq n_5(5-1) = 24
        \]
        which means that $|H| = 30$. By \myref{problem-group-of-order-30-has-normal-subgroup-of-order-5} again, such a group has only a unique Sylow 5-subgroup.  Note $5 \nmid [G:H]$, so \myref{problem-normal-subgroup-of-G-contains-all-sylow-p-subgroups} implies all Sylow 5-subgroups of $G$ are in $H$. However, right at the start, we concluded that there are 6 Sylow 5-subgroups in $G$, so $H$ must have 6 Sylow 5-subgroups, a contradiction.
    \end{enumerate}
    Hence, $H$ does not exist, and so $G$ is simple.
\end{proof}

\begin{exercise}\label{exercise-group-of-order-12-has-normal-subgroup-of-3-or-4}
    Prove that a group of order 12 either has a normal subgroup of order 3, or a normal subgroup of order 4, or both.
\end{exercise}

\begin{exercise}\label{exercise-group-of-order-15-or-20-has-normal-subgroup-of-order-5}
    Prove that a group of each of the following orders has a normal subgroup of order 5.
    \begin{partquestions}{\alph*}
        \item 15
        \item 20
    \end{partquestions}
\end{exercise}

\begin{corollary}\label{corollary-A5-is-simple}
    The group $\An5$ is simple.
\end{corollary}
\begin{proof}
    \myref{exercise-A5-has-two-distinct-subgroups-of-order-5} (later) gives two distinct subgroups of order 5. Since $|\An{5}| = 60 = 2^2 \times 3 \times 5$, thus subgroups of order 5 are Sylow 5-subgroups. Therefore $\An5$ is simple by \myref{thrm-group-of-order-60-with->1-sylow-5-subgroup-is-simple}.
\end{proof}
\begin{exercise}\label{exercise-A5-has-two-distinct-subgroups-of-order-5}
    Consider the permutation $\sigma = \begin{pmatrix}1&3&2&4&5\end{pmatrix}$.
    \begin{partquestions}{\roman*}
        \item Explain why $\sigma \in \An{5}$.
        \item Find the order of the subgroup $\langle \sigma \rangle$.
        \item Find another subgroup of $\An{5}$ with order 5.
    \end{partquestions}
\end{exercise}

We also state and prove a fairly obvious proposition.

\begin{proposition}\label{prop-An-stabilizer-of-i-is-isomorphic-to-A(n-1)}
    Let the integer $n \geq 3$. Let the set $\{1, 2, 3, \dots, n\}$ be denoted by $\mathcal{N}_n$. Suppose $\An{n}$ acts on $\mathcal{N}_n$ naturally. Then $\Stab{\An{n}}{r} \cong \An{n-1}$.
\end{proposition}
\begin{proof}
    We note that elements of $\Stab{\An{n}}{r}$ are permutations that fix $r$, thereby permuting the $n - 1$ other elements. Therefore the elements of $\Stab{\An{n}}{r}$ are even permutations on $n - 1$ elements, i.e. $\Stab{\An{n}}{r} \cong \An{n-1}$.
\end{proof}

With these results, we are ready to prove the main result of this section.

\begin{theorem}\label{thrm-An-is-simple-for-n>=5}
    The group $\An{n}$ is simple if $n \geq 5$.
\end{theorem}
\begin{proof}[Proof (see {\cite[Theorem 4.24]{dummit_foote_2004}})]
    We induct on $n$. For brevity, let $\mathcal{N}_n = \{1, 2, 3, \dots, n\}$. The base case of $n = 5$ is covered by \myref{corollary-A5-is-simple}. Assume that $\An{k-1}$ is simple for some $k \geq 6$; we will prove that $\An{k}$ is also simple.

    Let $G = \An{k}$ and, seeking a contradiction, assume that $G$ has a non-trivial proper normal subgroup $H$. Let $G$ act on $\mathcal{N}_{k}$ naturally; thus we see that $\Stab{G}{i} \leq G$ with $\Stab{G}{i} \cong \An{k-1}$ (\myref{prop-An-stabilizer-of-i-is-isomorphic-to-A(n-1)}) for any $i \in \mathcal{N}_k$. Note $\An{k-1}$ is simple by the induction hypothesis, so $\Stab{G}{i}$ is simple for each $i \in \mathcal{N}_{k}$.

    Suppose first that there is some non-identity $\pi \in H$ such that $\pi(i) = i$ for some $i \in \mathcal{N}_{k}$. This means that $\pi$ fixes $i$; thus $\pi \in H \cap \Stab{G}{i}$. Note that since $H \lhd G$ and $\Stab{G}{i} \leq G$ thus $H \cap \Stab{G}{i} \lhd \Stab{G}{i}$ by the Second Isomorphism Theorem (\myref{thrm-isomorphism-2}), statement 4. But as $\Stab{G}{i}$ is simple (and non-trivial) we must have $H \cap \Stab{G}{i} = \Stab{G}{i}$. Therefore $\Stab{G}{i} \subseteq H$ which means $\Stab{G}{i} \leq H$. Now by \myref{exercise-conjugate-of-stabilizer} (later), for any $\sigma \in G$, we know that $\sigma\Stab{G}{i}\sigma^{-1} = \Stab{G}{\sigma(i)}$. Therefore we see
    \[
        \sigma\Stab{G}{i}\sigma^{-1} \leq \sigma H\sigma^{-1} = H
    \]
    since $H \lhd G$. Thus, for any $j \in \mathcal{N}_{k+1}$, there exists $\sigma \in G$ where $\sigma(i) = j$ such that
    \[
        \sigma\Stab{G}{i}\sigma^{-1} = \Stab{G}{j} \leq H.
    \]

    Note that any $\lambda \in G$ may be written as a product of an even number of transpositions (\myref{thrm-parity-of-permutation}), say $2t$ transpositions. Thus, we may write $\lambda = \lambda_1\lambda_2\cdots\lambda_t$ where each $\lambda_i$ is a product of two transpositions. Now as $k \geq 5$, each $\lambda_i$ (which could at most consist of two disjoint cycles of 4 elements) must fix at least one element in $\mathcal{N}_{k}$, say $j$. That is, $\lambda_i \in \Stab{G}{j}$ for some $j \in \mathcal{N}_{k}$. Since $\lambda_i \in \Stab{G}{j} \leq H$ for some $j \in \mathcal{N}_{k}$, thus $\lambda_i \in H$. Hence $\lambda = \lambda_1\lambda_2\cdots\lambda_t \in H$. Therefore, any element in $G$ is also in $H$, meaning $G \subseteq H$, contradicting the fact that $H \lhd G$.

    We conclude that for any $\pi \in H$, if $\pi \neq \id$ then $\pi(i) \neq i$ for all $i \in \mathcal{N}_k$. The contrapositive of this statement is that if $\pi(i) = i$ for some $i \in \mathcal{N}_k$ then $\pi = \id$. Now suppose $\pi_1, \pi_2 \in H$ and $\pi_1(i) = \pi_2(i)$ for some $i \in \mathcal{N}_{k}$. Then $\pi_2^{-1}\pi_1(i) = i$, which implies $\pi_2^{-1}\pi_1 = \id$. Hence $\pi_1 = \pi_2$. Therefore, if $\pi_1, \pi_2 \in H$ and $\pi_1(i) = \pi_2(i)$ for some $i \in \mathcal{N}_{k}$, then $\pi_1 = \pi_2$.

    Now suppose a non-identity $\pi_1 \in H$ exists such that the cycle decomposition of $\pi_1$ contains a cycle of length of at least 3, say
    \[
        \pi_1 = \begin{pmatrix}a_1&a_2&a_3&\cdots\end{pmatrix} \begin{pmatrix}b_1&b_2&\cdots\end{pmatrix}\cdots
    \]
    where $a_1$, $a_2$, $a_3$, $b_1$, $b_2$, etc. are distinct (which is possible since $k \geq 5$). We note an element $\sigma \in G$ exists such that $\sigma(a_1) = a_1$, $\sigma(a_2) = a_2$, but $\sigma(a_3) \neq a_3$ because $k \geq 4$ (for example, the permutation $\begin{pmatrix}a_3 & a_4\end{pmatrix}$). Then \myref{exercise-conjugation-of-permutation-by-another} (later) tells us that
    \[
        \sigma\pi_1\sigma^{-1} = \begin{pmatrix}a_1&a_2&\sigma(a_3)&\cdots\end{pmatrix} \begin{pmatrix}\sigma(b_1)&\sigma(b_2)&\cdots\end{pmatrix}\cdots.
    \]
    Set $\pi_2 = \sigma\pi_1\sigma^{-1}$, which is clearly distinct from $\pi_1$. Then we see $\pi_1(a_1) = \pi_2(a_1) = a_2$, contrary to the above observation that $\pi_1(i) = \pi_2(i)$ for any $i \in \mathcal{N}_k$ implies $\pi_1 = \pi_2$. Therefore only 2-cycles can appear in the cycle decomposition of non-identity elements of $H$.

    Let $\pi_1 \in H$ be a non-identity element, so that
    \[
        \pi_1 = \begin{pmatrix}a_1&a_2\end{pmatrix} \begin{pmatrix}a_3&a_4\end{pmatrix} \begin{pmatrix}a_5&a_6\end{pmatrix}\cdots
    \]
    where each $a_i$ is distinct (note that $k \geq 6$ is used above). Consider the permutation $\sigma = \begin{pmatrix}a_1&a_2\end{pmatrix} \begin{pmatrix}a_3&a_5\end{pmatrix}$, which is in $G$ since it is made up of 2 transpositions. Then \myref{exercise-conjugation-of-permutation-by-another} again gives
    \[
        \sigma\pi_1\sigma^{-1} = \begin{pmatrix}a_1&a_2\end{pmatrix} \begin{pmatrix}a_5&a_4\end{pmatrix} \begin{pmatrix}a_3&a_6\end{pmatrix}\cdots.
    \]
    Setting $\pi_2 = \sigma\pi_1\sigma^{-1}$ again gives two distinct permutations $\pi_1$ and $\pi_2$ where $\pi_1(a_1) = \pi_2(a_1) = a_2$, again contrary to the above observation.

    We conclude that such a non-trivial proper normal subgroup $H$ of $\An{k}$ cannot exist. Thus, $\An{k-1}$ being simple implies that $\An{k}$ is also simple.

    By mathematical induction, $\An{n}$ is simple for all $n \geq 5$.
\end{proof}

\begin{exercise}\label{exercise-conjugate-of-stabilizer}
    Let $S$ be a non-empty set and let $G \leq \Sym{S}$ act on $S$. Show that $\sigma\Stab{G}{x}\sigma^{-1} = \Stab{G}{\sigma(x)}$ for any $\sigma \in G$ and $x \in S$.
\end{exercise}

\begin{exercise}\label{exercise-conjugation-of-permutation-by-another}
    Let $\sigma, \pi \in \Sn{n}$. Suppose $\sigma$ has cycle decomposition
    \[
        \begin{pmatrix}a_1&a_2&\cdots&a_{k_1}\end{pmatrix} \begin{pmatrix}b_1&b_2&\cdots&b_{k_2}\end{pmatrix}\cdots,
    \]
    where $a_1, a_2, \dots, a_{k_1}, b_1, b_2, \dots, b_{k_2}, \dots$ are all distinct. Show that
    \[
        \pi\sigma\pi^{-1} = \begin{pmatrix}\pi(a_1)&\cdots&\pi(a_{k_1})\end{pmatrix} \begin{pmatrix}\pi(b_1)&\cdots&\pi(b_{k_2})\end{pmatrix}\cdots,
    \]
    that is, $\pi\sigma\pi^{-1}$ is obtained from $\sigma$ by replacing each entry $i$ by $\pi(i)$.
\end{exercise}

\begin{corollary}
    The group $\An{n}$ is simple for $n \geq 3$ and $n \neq 4$.
\end{corollary}
\begin{proof}
    We note $\An3$ has order $\frac{3!}{2} = 3$ which is prime, so $\An3 \cong \Cn3$ which is simple. Also $\An{n}$ is simple for $n \geq 5$ by \myref{thrm-An-is-simple-for-n>=5}.
\end{proof}

We note that $\An4$ is non-simple by the solution of \myref{problem-S4-composition-series}, in which we found that $\An4$ has a unique composition series of
\[
    1 \lhd \Cn2 \lhd \mathrm{V} \lhd \An4
\]
up to isomorphism.

\section{Groups of Lie Type}
We briefly mention groups of Lie type; we will not prove any significant results here.

Groups of Lie (pronounced ``lee'') type\index{groups of Lie type} usually refers to finite groups that are closely related to the group of rational points of a reductive linear algebraic group with values in a finite field. We will cover finite fields in part III. We briefly mention these groups here.

The list below, taken from \cite{wikipedia_list-of-simple-groups}, is a list of the families of simple groups of Lie type. In what follows, $n$ is a positive integer and $q$ is a positive power of a prime number $p$.
\begin{itemize}
    \item \term{Classical Chevalley groups}\index{Chevalley groups!classical}: there are 4 families of simple groups.
    \begin{itemize}
        \item $A_n(q)$, except for $A_1(2)$ and $A_1(3)$. There are several duplicates, which are
        \begin{itemize}
            \item $A_1(4) \cong A_1(5) \cong \An{5}$;
            \item $A_1(7) \cong A_2(2)$;
            \item $A_1(9) \cong \An{6}$; and
            \item $A_3(2) \cong \An{8}$.
        \end{itemize}
        We note that $\An{n}$ is not the same as $A_n(q)$. We distinguish between the alternating group of degree $n$ ($\An{n}$) and the groups of Lie type $A_n(q)$ by letting the latter be in italics and the former be in `normal' font.

        \item $B_n(q)$ for $n > 1$, except for $B_2(2)$. There are several duplicates, which are
        \begin{itemize}
            \item $B_n(2^m) \cong C_n(2^m)$; and
            \item $B_2(3) \cong {^2A_3(2^2)} = {^2A_3(4)}$, where ${^2A_3(4)}$ is a classical Steinberg group.
        \end{itemize}
        \item $C_n(q)$ for $n > 2$. The only duplicate is $C_n(2^m) \cong B_n(2^m)$ mentioned earlier.
        \item $D_n(q)$ for $n > 3$.
    \end{itemize}

    \item \term{Exceptional Chevalley groups}\index{Chevalley groups!exceptional}: there are 5 families of such groups.
    \begin{itemize}
        \item $E_6(q)$;
        \item $E_7(q)$;
        \item $E_8(q)$;
        \item $F_4(q)$; and
        \item $G_2(q)$, except for $G_2(2)$.
    \end{itemize}

    \item \term{Classical Steinberg groups}\index{Steinberg groups!classical}: there are 2 families of simple groups.
    \begin{itemize}
        \item ${^2A_n(q^2)}$ for $n > 1$, except for ${^2A_2(2^2)} = {^2A_2(4)}$. The only duplicate is ${^2A_3(2^2)} \cong B_2(3)$ mentioned earlier.
        \item ${^2D_n(q^2)}$ for $n > 3$.
    \end{itemize}

    \item \term{Exceptional Steinberg groups}\index{Steinberg groups!exceptional}: there are 2 families of simple groups.
    \begin{itemize}
        \item ${^2E_6(q^2)}$; and
        \item ${^3D_4(q^3)}$.
    \end{itemize}

    \item \term{Suzuki groups}\index{Suzuki groups}: there is 1 family of simple groups, which is ${^2B_2(q)}$ where $q = 2^{2n+1}$ and $n \geq 1$. Such a group has order $q^2(q^2+1)(q-1)$.

    \item \term{Ree groups}\index{Ree groups}: there are 2 families of simple groups.
    \begin{itemize}
        \item $^2F_4(q)$ where $q = 2^{2n+1}$ and $n \geq 1$. The order of such a group is $q^{12}(q^6+1)(q^4-1)(q^3+1)(q-1)$.
        \item $^2G_2(q)$ where $q = 3^{2n+1}$ and $n \geq 1$. The order of such a group is $q^3(q^3+1)(q-1)$.
    \end{itemize}
\end{itemize}

There is also the \term{Tits group}\index{Tits group}, $^2F_4(2)'$, with an order of $17,971,200$. It is the commutator subgroup of $^2F_4(2)$, which is a Lie group but not a simple group. The fact that $^2F_4(2)'$ is linked to Ree groups makes most authors consider it not a sporadic group (see below).

\section{The Sporadic Groups}
Along with the 18 infinite families of simple groups, there are also 26 sporadic simple groups\index{sporadic group} that do not fall within the families (27 if the Tits group is considered a sporadic group).

We first list four categories of sporadic groups.
\begin{table}[h]
    \centering
    \begin{tabular}{|l|l|l|}
        \hline
        \textbf{Name} & \textbf{Symbol} & \textbf{Order} \\ \hline
        \multirow{5}{*}{\term{Mathieu Groups}\index{Mathieu groups}} & $\mathrm{M}_{11}$ & 7,290 \\ \cline{2-3}
        & $\mathrm{M}_{12}$ & 95,040 \\ \cline{2-3}
        & $\mathrm{M}_{22}$ & 443,520 \\ \cline{2-3}
        & $\mathrm{M}_{23}$ & 10,200,960 \\ \cline{2-3}
        & $\mathrm{M}_{24}$ & 244,823,040 \\ \hline
        \multirow{4}{*}{\term{Janko Groups}\index{Janko groups}} & $\mathrm{J}_1$ & 175,560 \\ \cline{2-3}
        & $\mathrm{J}_2$ & 604,800 \\ \cline{2-3}
        & $\mathrm{J}_3$ & 50,232,960 \\ \cline{2-3}
        & $\mathrm{J}_4$ & 86,775,571,046,077,562,880 \\ \hline
        \multirow{3}{*}{\term{Conway Groups}\index{Conway groups}} & $\mathrm{Co}_3$ & 495,766,656,000 \\ \cline{2-3}
        & $\mathrm{Co}_2$ & 42,305,421,312,000 \\ \cline{2-3}
        & $\mathrm{Co}_1$ & 4,157,776,806,543,360,000 \\ \hline
        \multirow{3}{*}{\term{Fischer Groups}\index{Fischer groups}} & $\mathrm{Fi}_{22}$ & 64,561,751,654,400 \\ \cline{2-3}
        & $\mathrm{Fi}_{23}$ & 4,089,470,473,293,004,800 \\ \cline{2-3}
        & $\mathrm{Fi}_{24}$ & 1,255,205,709,190,661,721,292,800 \\ \hline
    \end{tabular}
\end{table}



More sporadic groups are listed below.
\begin{table}[h]
    \centering
    \begin{tabular}{|l|l|l|}
        \hline
        \textbf{Group}        & \textbf{Symbol} & \textbf{Order}  \\ \hline
        \term{Higman-Sims group}\index{Higman-Sims group}     & HS              & 44,352,000      \\ \hline
        \term{McLaughlin group}\index{McLaughlin group}      & McL             & 898,128,000     \\ \hline
        \term{Held group}\index{Held group}            & He              & 4,030,387,200   \\ \hline
        \term{Rudvalis group}\index{Rudvalis group}        & Ru              & 145,926,144,000 \\ \hline
        \term{Suzuki sporadic group}\index{Suzuki sporadic group} & Suz             & 448,345,497,600 \\ \hline
        \term{O'Nan group}\index{O'Nan group}           & $\mathrm{O'N}$  & 460,815,505,920 \\ \hline
        \term{Harada-Norton group}\index{Harada-Norton group}   & HN              & 273,030,912,000,000    \\ \hline
        \term{Lyons group}\index{Lyons group}           & Ly              & 51,765,179,004,000,000 \\ \hline
        \term{Thompson group}\index{Thompson group}        & Th              & 90,745,943,887,872,000 \\ \hline
    \end{tabular}
\end{table}

The remaining 2 sporadic groups are special in that they have extremely large order.
\begin{itemize}
    \item The \term{Baby Monster group}\index{Baby Monster group}, usually denoted $\mathrm{B}$, has order
    \begin{align*}
        &2^{41} \times 3^{13} \times 5^6 \times 7^2 \times 11 \times 13 \times 17 \times 19 \times 23 \times 31 \times 47\\
        &= 4,154,781,481,226,426,191,177,580,544,000,000.
    \end{align*}
    \item The \term{Monster group}\index{Monster group}, usually denoted $\mathrm{M}$, has order $2^{46} \times 3^{20} \times 5^9 \times 7^6 \times 11^{2} \times 13^3 \times 17 \times 19 \times 23 \times 29 \times 31 \times 41 \times 47 \times 59 \times 71$ which equals 808,017,424,794,512,875,886,459,904,961,710,757,005,754,368,\linebreak000,000,000. It is the largest sporadic group.
\end{itemize}

\section{The Classification Theorem of Finite Simple Groups}
One might rightly wonder what the importance of listing out all of these different types of simple groups are. It turns out that, amazingly, that these results provide a complete classification of what a finite simple group can really be. This is captured in the Classification Theorem of Finite Simple Groups\index{Classification Theorem of Finite Simple Groups}, which is sometimes called the enormous theorem\index{Enormous Theorem}.

\begin{theorem}[Classification Theorem]
    Every finite simple group is isomorphic to either
    \begin{itemize}
        \item a cyclic group of prime order;
        \item an alternating group with degree of at least 5;
        \item a group in the 16 infinite families of groups of Lie type, or the Tits group; or
        \item one of 26 sporadic groups.
    \end{itemize}
\end{theorem}

The proof of this theorem required tens of thousands of pages in hundreds of articles, written by a large number of authors that were  published mostly between 1955 to 2004. The longest paper, and the last paper needed to fill in the gap for quasithin groups, was published in 2004 by Aschbacher and Smith and spanned in a 1221 pages. But after all that work, mathematicians had a complete classification of all finite simple groups.
