\chapter{Subgroups}
We introduced the simple properties of groups and simple examples of groups in the previous chapter. Sometimes, subsets of a group can also be a group. Such groups are called subgroups; we explore important properties and results relating to them in this chapter, with the most important result being Lagrange's theorem.

\section{Definition and Examples}
We look at the definition of a subgroup of a group.
\begin{definition}
    Let $G$ be a group with operation $\ast$. Let $H$ be a subset of $G$. Then $H$ is said to be a \textbf{subgroup}\index{subgroup} of $G$ if it satisfies the following conditions.
    \begin{enumerate}
        \item \textbf{Closure}: For all $x$ and $y$ in $H$, we have $x \ast y$ is also in $H$.
        \item \textbf{Identity}: The identity of the group $G$ is in $H$.
        \item \textbf{Inverse}: For all elements $x$ in $H$, there exists an element $x^{-1}$ in $H$ such that $x \ast x^{-1} = x^{-1} \ast x = e$.
    \end{enumerate}
\end{definition}
\begin{remark}
    Equivalently, $H$ is a subgroup of $G$ if $H$ is a subset of $G$ and $H$ is a group under $\ast$.
\end{remark}
We write $H \leq G$ if $H$ is a subgroup of $G$.

\begin{example}\label{example-subgroups-of-Z}
    Let's look at all possible subgroups of the group $(\mathbb{Z}, +)$. Suppose $H$ is a subgroup of $G$ with $H \neq \{0\}$.

    Let $n$ be the smallest positive integer in $H$. Let $m$ be any other number in $H$. Then by Euclid's division lemma (\myref{lemma-euclid-division}), we may write $m = nq + r$ where $q$ and $r$ are integers such that $0 \leq r < n$. Hence,
    \[
        r = m + \underbrace{(-n) +(-n) +(-n) + \cdots + (-n)}_{q\text{ times}}.
    \]
    Note that since both $m$ and $-n$ are in $H$, thus $r = m + (-n) + \cdots + (-n)$ is also in $H$, since $H$ is closed under addition. But $0 \leq r < n$ and $n$ is the \textit{smallest} positive integer in $H$. Thus, $r \not> 0$ which means $r = 0$. Therefore $m = nq$, i.e. every element in $H$ is a multiple of the smallest positive integer in $H$.

    Hence,
    \[
    H = \{nk \vert k \in \mathbb{Z}\},
    \]
    which is often written as the group $n\mathbb{Z}$.
\end{example}

\begin{exercise}
    Let $G$ be a group with identity $e$. Prove that $\{e\} \leq G$.
\end{exercise}

We note that the trivial group (i.e., $\{e\}$) is always a subgroup of any group. This is known as the \textbf{trivial subgroup}\index{subgroup!trivial}. We also note that the group itself is a subgroup of itself (that is, $G \leq G$). Any subgroup that are \textbf{not} these two groups is known as a \textbf{proper subgroup}\index{subgroup!proper} of $G$, and we write $H < G$ if $H$ is a proper subgroup of $G$.

\section{Subgroup Test}
To prove that a subset of a group is a subgroup using the axioms is too tedious. Wouldn't it be nice if we have a simple test to determine if a subset is a subgroup? Well, there is; it is called the \textbf{subgroup test}\index{subgroup test}.
\begin{theorem}[Subgroup Test]\label{thrm-subgroup-test}
    Let $G$ be a group and $H \subseteq G$ such that $H \neq \emptyset$ (i.e., $H$ is not the empty set). Then $H \leq G$ if and only if $xy^{-1} \in H$ for all $x, y \in H$.
\end{theorem}
\begin{proof}
    We prove the forward direction first. Suppose $H \leq G$ and let $x$ and $y$ be elements in $H$. Since $H$ is a subgroup of $G$, inverses exist. Thus $y^{-1}$ is in $H$. Also, since $H$ is a subgroup of $G$, thus $H$ is closed under the group operation. Hence, $xy^{-1}$ is in $H$. Therefore, if $H \leq G$ then $xy^{-1} \in H$ for all $x, y \in H$.

    We now prove the reverse direction. Suppose $xy^{-1}\in H$ for all $x$ and $y$ in $H$.
    \begin{itemize}
        \item Suppose $h$ is an arbitrary element in $H$. Set $x = h$ and $y = h$. Then $xy^{-1} = hh^{-1} = e$ is in $H$. Thus the identity of $G$ is present in $H$.
        \item Set $x = e$ and let $y = h$. Then $xy^{-1} = eh^{-1} = h^{-1}$ is in $H$. Hence every element in $H$ has an inverse in $H$.
        \item Let $a$ and $b$ be elements in $H$. Then by above point, $b^{-1}$ is in $H$. Set $x = a$ and $y = b^{-1}$. Then $xy^{-1} = a\left(b^{-1}\right)^{-1} = ab$ is in $H$. Hence $H$ is closed under the group operation.
    \end{itemize}
    Therefore if $xy^{-1}$ is in $H$ if for all $x$ and $y$ in $H$, then $H \leq G$.

    This completes the proof of the subgroup test.
\end{proof}

\begin{remark}
We usually verify the `non-empty' and `subset' requirement by checking if the identity of $G$ is in $H$.
\end{remark}

We look at some examples of the use of the subgroup test.
\begin{example}\label{example-centralizer-of-a-subset}
    Let $G$ be a group and $S \subseteq G$ where $S \neq \emptyset$. The \textbf{centralizer}\index{centralizer} of $S$ in $G$ is the set
    \[
        \C{G}{S} = \left\{g \in G \vert \text{for all } s \in S \text{ we have } gs = sg\right\}.
    \]
    We claim that $\C{G}{S}$ is a subgroup of $G$.
    \begin{proof}
        We show that for all elements $x$ and $y$ in $\C{G}{S}$, $xy^{-1}$ is also in the centralizer. Note that the condition $gs = sg$ is equivalent to the condition $sg^{-1} = g^{-1}s$.

        Clearly for all $s \in S$ we have $es = se$, so $e \in \C{G}{S}$. This means that $\C{G}{S}$ is non-empty and $\C{G}{S}$ is a subset of $G$.
        
        Now suppose $x$ and $y$ are in $\C{G}{S}$. Let $s$ be in $S$. Then
        \begin{align*}
            (xy^{-1})s &= x(y^{-1}s) & (\text{associativity})\\
            &= x(sy^{-1}) & (\text{since } g^{-1}s = sg^{-1})\\
            &= (xs)y^{-1} & (\text{associativity})\\
            &= (sx)y^{-1} & (\text{since } gs = sg)\\
            &= s(xy^{-1}) & (\text{associativity})
        \end{align*}
        so $(xy^{-1})s = s(xy^{-1})$ for all $s$ in $S$. Thus, $xy^{-1}$ is in $\C{G}{S}$ for all $x$ and $y$ in $\C{G}{S}$, so by subgroup test $\C{G}{S} \leq G$.
    \end{proof}
\end{example}
\begin{remark}
    In the case where the set $S$ contains a single element, say $S = \{x\}$, we write $\C{G}{x}$.
\end{remark}

\begin{example}\label{example-center-of-group}
    Let $G$ be a group. The \textbf{center}\index{center} of a group $G$ is given by the set
    \[
        \Z{G} = \{z \in G \vert gz = zg \text{ for all } g \in G\}.
    \]
    We claim that $\Z{G}$ is a subgroup of $G$.

    \begin{proof}
	    Note that $e \in \Z{G}$ since $ge = eg$ for all $g \in G$.

	    Let $x$ and $y$ be in $\Z{G}$, meaning that $gx = xg$ and $gy = yg$ for all $g \in G$. Note that $gy = yg$ implies $y^{-1}g = gy^{-1}$, so $y^{-1} \in \Z{G}$. Now,
	    \begin{align*}
	        (xy^{-1})g &= x(y^{-1}g)\\
	        &= x(gy^{-1}) & (\text{since } y^{-1} \in \Z{G})\\
	        &= (xg)y^{-1} & (\text{associativity})\\
	        &= (gx)y^{-1} & (\text{since } x \in \Z{G})\\
	        &= g(xy^{-1}) & (\text{associativity})
	    \end{align*}
	    which means that $xy^{-1} \in \Z{G}$. Hence $\Z{G} \leq G$ by the subgroup test.
    \end{proof}
\end{example}

\begin{exercise}\label{exercise-conjugate-subgroup}
    Let $G$ be a group, $H \leq G$, and $g \in G$. Define the set
    \[
        S = \{ghg^{-1} \vert h \text{ is in } H\}.
    \]
    Prove that $S \leq G$ under the group operation of $G$.
\end{exercise}

\newpage

\section{Cosets}
We now introduce the idea of \textbf{cosets}\index{coset} of a group $G$.
\begin{definition}
    Let $G$ be a group, $H \leq G$, and $g$ be an element of $G$. Then
    \begin{itemize}
        \item the \textbf{left coset}\index{coset!left} of $H$ in $G$ by $g$ is $gH = \{gh \vert h \in H\}$; and
        \item the \textbf{right coset}\index{coset!right} of $H$ in $G$ by $g$ is $Hg = \{hg \vert h \in H\}$.
    \end{itemize}
\end{definition}

\begin{remark}
    The subgroup $H$ is both a left and right coset of $H$ in $G$. This is because the element $g$ in question is $g = e$, so $gH = eH = H$ and $Hg = He = H$.
\end{remark}

\begin{example}
    Let $G = D_3$, and $H = \{e, s\} \leq G$. The distinct left cosets of $H$ in $G$ are
    \begin{itemize}
        \item $eH = \{e, s\} = H$;
        \item $rH = \{r, rs\}$; and
        \item $r^2H = \{r^2, r^2s\}$.
    \end{itemize}
    Since each element of $G$ appears in one of these cosets, generating more using any other element would not give new cosets. This is because a new coset would have to contain an element in common with one of the above cosets and therefore be identical to it. For example, $rsH = \{rs, rss\} = \{rs, r\} = rH$.

    The distinct right cosets of $H$ in $G$ are
    \begin{itemize}
        \item $He = \{e, s\} = H$;
        \item $Hr = \{r, sr\} = \{r, r^2s\}$; and
        \item $Hr^2 = \{r^2, sr^2\} = \{r^2, rs\}$.
    \end{itemize}
    Thus, for $D_3$, no left coset is a right coset, except for $H = eH = He$.
\end{example}

\begin{exercise}
    Let $G$ be the group $(\mathbb{Z}_8, \oplus_8)$. Let $H = \{0, 4\} \leq G$.
    \begin{partquestions}{\alph*}
        \item Explain why any left coset of $H$ in $G$ by an element $g$ is the same as the right coset of $H$ in $G$ by $g$.
        \item Find all distinct left cosets of $H$ in $G$.
    \end{partquestions}
\end{exercise}

We now state and prove a result that relates the equality of cosets.
\begin{lemma}[Coset Equality]\label{lemma-coset-equality}\index{Coset Equality}
    Let $G$ be a group, $H \leq G$, and $g_1$ and $g_2$ be elements in $G$. Then the following statements are equivalent.
    \begin{enumerate}[label=$(\arabic*)$]
        \item $g_1H = g_2H$
        \item $Hg_1^{-1} = Hg_2^{-1}$
        \item $g_1H \subseteq g_2H$
        \item $g_2 \in g_1H$
        \item $g_1^{-1}g_2 \in H$
    \end{enumerate}
\end{lemma}
\begin{proof}
    We prove the statements in order.

    \begin{itemize}
        \item $\boxed{(1) \implies (2)}$ Suppose $g_1H = g_2H$. We show $Hg_1^{-1} = Hg_2^{-1}$.

        Let $x \in Hg_1^{-1}$. Then $x = hg_1^{-1}$ for some $h$ in $H$. Thus $x^{-1} = \left(hg_1^{-1}\right)^{-1} = g_1h^{-1}$ by Shoes and Socks. Since $h^{-1}$ is in $H$, thus $x^{-1} = g_1h^{-1}$ is in $g_1H$.

        Since $g_1H = g_2H$, thus $x^{-1} \in g_2H$. Write $x^{-1} = g_2\hat{h}$ for some $\hat{h}$ in $H$. Thus $x = (g_2\hat{h})^{-1} = \hat{h}^{-1}g_2^{-1}$ by Shoes and Socks. Since $\hat{h}^{-1}$ is in $H$ thus $x = \hat{h}^{-1}g_2^{-1}$ is in $Hg_2^{-1}$.

        Hence, any element $x \in Hg_1^{-1}$ is also in $Hg_2^{-1}$, i.e. $Hg_1^{-1} \subseteq Hg_2^{-1}$. A similar argument shows that $Hg_2^{-1} \subseteq Hg_1^{-1}$. Thus $Hg_1^{-1} = Hg_2^{-1}$ as required.

        \item $\boxed{(2) \implies (3)}$ Suppose $Hg_1^{-1} = Hg_2^{-1}$ and take $x \in g_1H$. Thus $x = g_1h$ for some $h$ in $H$. Therefore $x^{-1} = (g_1h)^{-1} = h^{-1}g_1^{-1} \in Hg_1^{-1}$ since $h^{-1}$ is in $H$. By assumption $Hg_1^{-1} = Hg_2^{-1}$ so $x^{-1} \in Hg_2^{-1}$. Let $x^{-1} = \hat{h}g_2^{-1}$ for some $\hat{h}$ in $H$. Then $x = \left(\hat{h}g_2^{-1}\right)^{-1} = g_2\hat{h}^{-1} \in g_2H$ since $\hat{h}^{-1}$ is in $H$. Hence $x$ is in $g_2H$. Therefore, for any $x \in g_1H$, $x$ will also be in $g_2H$. Thus $g_1H \subseteq g_2H$.

        \item $\boxed{(3) \implies (4)}$
        Suppose $g_1H \subseteq g_2H$. Then for all $x$ in $g_1H$, $x$ is also in $g_2H$. Note that $g_1 = g_1e$. Since $e$ is in $H$ (as $H \leq G$) thus $g_1e \in g_1H \subseteq g_2H$ by assumption. Thus $g_1 \in g_2H$ as needed.

        \item $\boxed{(4) \implies (5)}$
        Suppose $g_1 \in g_2H$. Then $g_1 = g_2h$ for some $h$ in $H$. Thus,
        \begin{align*}
            &g_1^{-1}g_1 = g_1^{-1}g_2h & (\text{left multiply by } g_1^{-1})\\
            &e = g_1^{-1}g_2h\\
            &h^{-1} = g_1^{-1}g_2hh^{-1} & (\text{right multiply by } h^{-1})\\
            &h^{-1} = g_1^{-1}g_2.
        \end{align*}
        Since $h^{-1}$ is an element in $H$ thus $g_1^{-1}g_2 = h^{-1}$ is also in $H$, meaning $g_1^{-1}g_2 \in H$.

        \newpage

        \item $\boxed{(5) \implies (1)}$
        Suppose $g_1^{-1}g_2$ is an element of $H$. Thus, $g_1^{-1}g_2 = \hat{h}$ for some $\hat{h}$ in $H$. Let $x$ be an element from $g_1H$, so $x = g_1h$ for some $h$ in $H$. Then
        \begin{align*}
            x^{-1} &= (g_1h)^{-1}\\
            &= h^{-1}g_1^{-1} & (\text{Shoes and Socks})\\
            &= h_1^{-1}g_1^{-1}(g_2g_2^{-1}) & (\text{since }g_2g_2^{-1} = e)\\
            &= h_1{-1}(g_1^{-1}g_2)g_2^{-1}\\
            &= h^{-1}\hat{h}g_2^{-1}
        \end{align*}
        which means $x = \left(h^{-1}\hat{h}g_2^{-1}\right)^{-1} = g_2\hat{h}^{-1}h$, which is an element of $g_2H$ since $\hat{h}^{-1}h$ is in $H$.

        Thus, for all $x$ in $g_1H$, $x$ is also in $g_2H$ which means $g_1H \subseteq g_2H$. A similar argument can be used to show that $g_2H \subseteq g_1H$. Hence $g_1H = g_2H$.
    \end{itemize}

    Thus, $(1) \implies (2) \implies (3) \implies (4) \implies (5) \implies (1)$, completing the proof.
\end{proof}

\begin{exercise}\label{exercise-intersection-of-cosets}
    Let $G$ be a group, $H \leq G$, and $g_1$ and $g_2$ be elements in $G$. Prove that if $g_1H \cap g_2H \neq \emptyset$ then $g_1H = g_2H$.
\end{exercise}

\newpage

\section{Lagrange's Theorem}
Lagrange's theorem is an important result relating the order of a subgroup and the order of the group itself. Before that, we introduce the idea of the \textbf{index} of a subgroup.
\begin{definition}
    Let $G$ be a group and $H \leq G$. The \textbf{index}\index{index} of $H$ in $G$, denoted by $[G : H]$, is the number of left cosets of $H$ in $G$.
\end{definition}
Note that since the number of left cosets is the number of right cosets, $[G : H]$ can be defined as the number of cosets of $H$ in $G$.

\begin{example}
    If $G = D_3$ and $H = \{e, s\}$, then there are 3 distinct left cosets, namely $H$, $rH$, and $r^2H$. Thus, $[G:H] = 3$ in this case.
\end{example}

We require two lemmas to properly prove Lagrange's theorem.
\begin{lemma}\label{lemma-left-coset-partition}
    Let $G$ be a group and $H \leq G$. Then distinct left cosets of $H$ in $G$ partition $G$.
\end{lemma}
\begin{proof}
    To prove that the distinct left cosets of $H$ in $G$ partition $G$, we need to show two things.
    \begin{itemize}
        \item The intersection of any 2 left cosets is either the empty set or is one of the left cosets (i.e., distinct left cosets are disjoint).
        \item The union of all left cosets is the group.
    \end{itemize}
    The first bullet point is proven by \myref{exercise-intersection-of-cosets}, so we work only on the second bullet point.

    Suppose $g$ is in $G$. We will find a left coset that $g$ is in. Clearly $g = ge$, and since $e$ is an element of $H$, thus $g = ge$ is an element of the left coset $gH$. So any element in $G$ belongs to a left coset of $H$ in $G$. Thus the union of all left cosets is the group.

    Hence, distinct left cosets of $H$ in $G$ partition $G$.
\end{proof}

\begin{lemma}\label{lemma-order-of-coset}
    Let $G$ be a group and $H \leq G$. Then $|H| = |gH|$ for all $g$ in $G$.
\end{lemma}
\begin{proof}
    Define the map $\phi: H \to gH$ such that $\phi(h) = gh$. To prove that $|H| = |gH|$, we need to show that $\phi$ is a bijection.
    \begin{itemize}
        \item \textbf{Injective}: Let $h$ and $\hat{h}$ be elements in $H$ such that $\phi(h) = \phi(\hat{h})$. Then $gh = g\hat{h}$ by definition of $\phi$. By cancellation law, we have $h = \hat{h}$ which means $\phi$ is injective.
        \item \textbf{Surjective}: Let $x$ be in $gH$. Thus $x = gh$ for some $h$ in $H$. Clearly $\phi(h) = gh = x$ so a pre-image of $x$ exists in $H$. Therefore $\phi$ is surjective.
    \end{itemize}
    Therefore $\phi$ is a bijection, which means $|H| = |gH|$.
\end{proof}

We are now ready to state and prove Lagrange's theorem.
\begin{theorem}[Lagrange]\label{thrm-lagrange}\index{Lagrange's Theorem}
    Let $G$ be a group and $H \leq G$. Then $|G| = [G:H]|H|$.
\end{theorem}
\begin{proof}
    Suppose $|G| = n$. Let $\mathcal{S} = \{g_1H, g_2H, g_3H, \dots, g_kH\}$ be the set containing all distinct left cosets of $H$ in $G$. Thus $k = [G: H]$.

    Note that distinct left cosets partition $G$ (\myref{lemma-left-coset-partition}). Hence
    \[
        G = \bigcup_{i=1}^k g_iH = g_1H \cup g_2H \cup \cdots \cup g_kH
    \]
    with $g_iH \cap g_jH = \emptyset$ if $i \neq j$. This means
    \begin{align*}
        |G| &= \sum_{i=1}^k|g_iH|\\
        &=\sum_{i=1}^k|H| & (\myref{lemma-order-of-coset})\\
        &= k|H|.
    \end{align*}
    Recall that we set $k = [G: H]$, therefore $|G| = [G:H]|H|$, proving Lagrange's theorem.
\end{proof}

\begin{exercise}
    Let $G$ be the group $(\mathbb{Z}_{99}, \oplus_{99})$. It is given that $H = \{0, 33, 66\}$ is a subgroup of $G$. What is the index of $H$ in $G$?
\end{exercise}

Let's look at some corollaries of Lagrange's theorem.
\begin{corollary}\label{corollary-order-of-group-multiple-of-order-of-element}
    Let $G$ be a finite group and let $g$ be an element in $G$. Then $|G|$ is a multiple of $|g|$.
\end{corollary}
\begin{proof}
    We show that $|G| = m|g|$ for some positive integer $m$. Clearly $|e| = 1$ which means $|G| = |G| \times 1 = |G||e|$, so suppose $g$ is a non-identity element of $G$ with order $n$. Consider $\mathcal{S} = \langle g \rangle = \{g, g^2, g^3, \dots, g^n\}$. Note that $\mathcal{S}$ is a (cyclic) subgroup of $G$ and $|\mathcal{S}| = n$. By Lagrange's theorem (\myref{thrm-lagrange}), $|G| = [G:\mathcal{S}]|\mathcal{S}| = [G:\mathcal{S}]|g|$. Hence, in this case, $m = [G:\mathcal{S}]$, proving the claim.
\end{proof}

\newpage

\begin{corollary}\label{corollary-group-with-prime-order-subgroups}
    A finite group $G$ with prime order $p$ has no proper subgroups.
\end{corollary}
\begin{proof}
    By Lagrange's theorem (\myref{thrm-lagrange}), the order of a subgroup must be a factor of the order of the group. Since the order of the group is prime, it only has 2 factors, namely 1 and $p$. The subgroup of order 1 is the trivial subgroup and the subgroup of order $p$ is $G$ itself. Hence $G$ has no proper subgroups.
\end{proof}

\begin{exercise}\label{exercise-prime-order-element}
    Let $G$ be a finite group with prime order $p$. Let $x$ be a non-identity element in $G$. Prove that $|x| = p$.
\end{exercise}

\begin{corollary}\label{corollary-group-with-prime-order-is-cyclic}
    A finite group $G$ with prime order $p$ is cyclic.
\end{corollary}
\begin{proof}
    By \myref{exercise-prime-order-element}, any non-identity element $g$ in $G$ has $|g| = p$. Thus $\langle g \rangle \neq \{e\}$.

    Note that $\langle g \rangle = \{g, g^2, g^3, \dots, g^p\} \leq G$. However, by \myref{corollary-group-with-prime-order-subgroups}, the only subgroups of $G$ are $\{e\}$ and $G$. Since $\langle g \rangle \neq \{e\}$ thus $\langle g \rangle = G$, meaning $G$ is cyclic with generator $g$.
\end{proof}

\section{Normal Subgroups}
We now look at a special type of subgroup, known as a \textbf{normal subgroup}.
\begin{definition}
    Let $G$ be a group and $N \leq G$. We say that $N$ is a \textbf{normal subgroup}\index{subgroup!normal} of $G$ if $gN = Ng$ for all $g \in G$.
\end{definition}
If $N$ is a normal subgroup of $G$, we write $N \unlhd G$. Furthermore if $N$ is a \textit{proper} normal subgroup of $G$, we write $N \lhd G$.

\newpage

Note that $gN = Ng$ is equivalent to the following two statements:
\begin{itemize}
    \item $gNg^{-1} = N$ for all $g \in G$. (One may interpret $gNg^{-1}$ as either the left coset $g(Ng^{-1})$ or the right coset $(gN)g^{-1}$.)
    \item $gng^{-1} \in N$ for all $g \in G$ and $n \in N$.
\end{itemize}

\begin{proposition}\label{prop-subgroup-of-abelian-group-is-normal}
    Any subgroup of an abelian group is normal.
\end{proposition}
\begin{proof}
    Let $G$ be an abelian group and $H \leq G$. Let $g$ be in $G$ and $h$ be in $H$. Then
    \begin{align*}
        ghg^{-1} &= g(hg^{-1})\\
        &= g(g^{-1}h) & (G \text{ is abelian})\\
        &= (gg^{-1})h & (\text{associativity})\\
        &= eh\\
        &= h
    \end{align*}
    which is an element of $H$. Thus $ghg^{-1}$ is an element of $H$ for all $g$ in $G$ and $h$ in $H$, meaning $H \unlhd G$.
\end{proof}

\begin{example}\label{example-normal-subgroups-of-d3}
    Let's find the normal subgroups of the dihedral group of order 6, $D_3$.

    Recall $D_3 = \{e, r, r^2, s, rs, r^2s\}$. Note that $|r| = 3$, $|s| = 2$, $\langle r \rangle = \{e, r, r^2\}$, and $\langle s \rangle = \{e, s\}$. We will show that $\langle r \rangle$ is a (proper) normal subgroup of $D_3$ but not $\langle s \rangle$. Note that since $r$ and $s$ are generators, we just need to check $s\langle r\rangle$, $\langle r\rangle s$, $r\langle s\rangle$, and $\langle s\rangle r$ to verify normality.

    \newpage

    For $\langle r \rangle$,
    \begin{itemize}
        \item $s\langle r\rangle = \{s, sr, sr^2\} = \{s, r^2s, rs\}$; and
        \item $\langle r\rangle s = \{s, rs, r^2s\}$,
    \end{itemize}
    so $s\langle r\rangle = \langle r \rangle s$ which means $\langle r \rangle \lhd D_3$.

    For $\langle s \rangle$,
    \begin{itemize}
        \item $r\langle s\rangle = \{r, rs\}$; and
        \item $\langle s \rangle r = \{r, sr\} = \{r, r^2s\}$,
    \end{itemize}
    and since $rs \neq r^2s$ thus $r\langle s\rangle \neq \langle s \rangle r$. Hence, $\langle s \rangle$ is not a normal subgroup of $D_3$.
\end{example}

\section{Quotient Groups}
We end this chapter by looking at a special (and useful) type of group: the quotient group. But before we can do that, we look at the idea of the set of left cosets.

\begin{definition}
    Let $G$ be a group and $H \leq G$. The \textbf{set of left cosets}\index{set of left cosets} is denoted by
    \[
        G/H = \{gH \vert g \in G \}.
    \]
\end{definition}
\begin{remark}
    We may sometimes write $G/H$ using fractions (i.e. $\frac GH$) if it serves to improve readability.
\end{remark}

Note that if $G/H$ is \textbf{not} a group, it is read ``$G$ by $H$''. Recall that the number of left cosets is $[G:H] = \frac{|G|}{|H|}$ by Lagrange's Theorem (\myref{thrm-lagrange}).

\begin{theorem}\label{thrm-quotient-group-requirement}
    Let $G$ be a group and $N \unlhd G$. Then $G / N$ forms a group called the \textbf{quotient group}\index{quotient group} with group operation $\star$ where
    \[
        (xN) \star (yN) = (xy)N.
    \]
\end{theorem}
For the quotient group, $G / N$ is read ``$G$ mod $N$''. Also note that, as per usual, we suppress the operation $\star$ and just write $(xN)(yN) = (xy)N$.
\begin{proof}
    Before we can prove that it forms a group, we need to show that $\star$ is well defined. This is because we worry that if $x_1N = x_2N$ and $y_1N = y_2N$ there may be a situation that $(x_1y_1)N \neq (x_2y_2)N$ under this operation. Thus we need to check if it is well defined before we can proceed.

    Suppose $x_1N = x_2N$ and $y_1N = y_2N$ where $x_1, x_2, y_1, y_2 \in G$. Then
    \begin{align*}
        (x_1N)(y_1N) &= (x_1y_1)N & (\text{by definition})\\
        &= x_1(y_1N) & (\text{left coset of } y_1N \text{ in } G \text{ by } x_1)\\
        &= x_1(y_2N) & (y_1N = y_2N)\\
        &= x_1(Ny_2) & (N \text{ is normal, so } y_2N=Ny_2)\\
        &= (x_1N)y_2 & (\text{right coset of } x_1N \text{ in } G \text{ by } y_2)\\
        &= (x_2N)y_2 & (x_1N = x_2N)\\
        &= x_2(Ny_2) & (\text{left coset of } Ny_2 \text{ in } G \text{ by } x_2)\\
        &= x_2(y_2N) & (N \text{ is normal, so } y_2N=Ny_2)\\
        &= (x_2y_2)N & (\text{left coset of } N \text{ in } G \text{ by } x_2y_2)\\
        &= (x_2N)(y_2N) & \text{(by definition)}
    \end{align*}
    so if $x_1N = x_2N$ and $y_1N = y_2N$ then we see that $(x_1N)(y_1N) = (x_2N)(y_2N)$, meaning that $\star$ is well defined.

    We can now show that $G/N$ under $\star$ satisfies the group axioms.
    \begin{enumerate}
        \item \textbf{Closure}: Assume $xN$ and $yN$ are in the set $G/N$. Then $(xN)(yN) = (xy)N$. Since $xy$ is in $G$ thus $(xy)N$ is in $G/N$, meaning that $G/N$ is closed under $\star$.
        \item \textbf{Associativity}: Take $xN$, $yN$, and $zN$ from $G/N$. Then
        \begin{align*}
            (xN)\left((yN)(zN)\right) &= (xN)\left((yz)N\right)\\
            &= (xyz)N\\
            &= \left((xy)z\right)N\\
            &= \left((xy)N\right)(zN)\\
            &= \left((xN)(yN)\right)(zN)
        \end{align*}
        so $\star$ is associative.
        \item \textbf{Identity}: Observe that $e$ is in $G$ so $eN = N$ is in $G / N$. Note that
        \[
        (eN)(xN) = (ex)N = xN \text{ and } (xN)(eN) = (xe)N = xN
        \]
        for any $x$ in $G$, so $eN = N$ is the identity in $G/N$.
        \item \textbf{Inverse}: Observe that for any $x$ in $G$, $x^{-1}$ is also in $G$, so
        \[
        (xN)(x^{-1}N) = (xx^{-1})N = eN = N
        \]
        and
        \[
        (x^{-1}N)(xN) = (x^{-1}x)N = eN = N,
        \]
        which means $(xN)^{-1}$ is $x^{-1}N$.
    \end{enumerate}
    Since the four group axioms are satisfied, thus $G/N$ is a group under the operation $\star$.
\end{proof}

\begin{example}
    Let's look at possible quotients of the group $D_3$, which has the underlying set of $\{e, r, r^2, s, rs, r^2s\}$. Recall from a previous example (\myref{example-normal-subgroups-of-d3}) that $\langle r \rangle \lhd D_3$. Thus $D_3 / \langle r \rangle$ is a quotient group, with $|D_3 / \langle r \rangle| = \frac{|D_3|}{|\langle r\rangle|} = \frac63 = 2$.

    Let's now look at the elements of $D_3 / \langle r \rangle$.
    \begin{align*}
        D_3 / \langle r \rangle  &= \{x\langle r \rangle \vert x \in D_3\}\\
        &= \left\{\{x, xr, xr^2\} \vert x \in D_3\right\}\\
        &= \{\{e, r, r^2\}, \{r, r^2, r^3\}, \{r^2, r^3, r^4\}, \\ &\quad\quad \{s, sr, sr^2\}, \{rs, rsr, rsr^2\}, \{r^2s, r^2sr, r^2sr^2\}\}\\
        &= \{\{e, r, r^2\}, \{r, r^2, e\}, \{r^2, e, r\}, \\ &\quad\quad \{s, sr, sr^2\}, \{sr^2, s, sr\}, \{sr, sr^2, s\}\}\\
        &= \left\{\{e, r, r^2\}, \{s, sr, sr^2\}\right\}\\
        &= \left\{\langle r\rangle, s\langle r \rangle\right\}
    \end{align*}
    Note also that $(s\langle r \rangle)^2 = s^2\langle r \rangle = \langle r\rangle$, so in fact $D_3 / \langle r \rangle$ has generator $s\langle r \rangle$, i.e. $D_3 / \langle r \rangle = \left\langle s\langle r \rangle \right\rangle$.
\end{example}

\begin{exercise}\label{exercise-quotient-group-of-cyclic-group-is-cyclic}
    Let $G$ be a finite cyclic group. Let $H$ be a subgroup of $G$.
    \begin{partquestions}{\roman*}
        \item Explain why $G/H$ is a quotient group.
        \item Show that $G/H$ is cyclic.
    \end{partquestions}    
\end{exercise}

\newpage

\section{Problems}
\begin{problem}
    Let $G = D_4$, the dihedral group of order 8. By considering the subgroup axioms, determine if the following are subgroups of $G$.
    \begin{partquestions}{\alph*}
        \item $\{e\}$
        \item $\{e, r, s\}$
        \item $\{r, r^2, r^3\}$
        \item $\{r, r^3, r^4, r^6\}$
    \end{partquestions}
\end{problem}

\begin{problem}
    Let $G$ be a group and $H \leq G$. Let
    \[
        K = \{x \in G \vert x^2 \in H\}.
    \]
    Prove the following statements.
    \begin{partquestions}{\alph*}
        \item $K \leq G$
        \item $H \leq K$
    \end{partquestions}
\end{problem}

\begin{problem}\label{problem-center-of-G}
    Let $G$ be a group.
    \begin{partquestions}{\alph*}
        \item Prove that $\Z{G}$ is a normal subgroup of $G$.
        \item Prove that $\Z{G} = G$ if and only if $G$ is abelian.
        \item Find the center of the group $D_4$.\newline
        (You may assume $|\Z{G}| < \frac12 |G|$. The reason becomes apparent with \myref{problem-quotient-of-group-mod-center-is-cyclic-implies-abelian}'s solution.)
    \end{partquestions}
\end{problem}

\newpage

\begin{problem}\label{problem-intersection-of-subgroups}
    Let $G$ be a group, and $H, K \leq G$. Prove or disprove the following statements.
    \begin{partquestions}{\alph*}
        \item $H \cap K \leq G$
        \item $H \cap K \leq H$
        \item $H \cup K \leq G$
        \item $H \cup K \leq H$
    \end{partquestions}
\end{problem}

\begin{problem}
    Let $G$ be a group of order 1024 and let $H$ be a proper subgroup of $G$. Determine the maximum order of $H$. Give an example of the groups $G$ and $H$ such that $H$ has this maximum order.
\end{problem}

\begin{problem}
    Let $G$ be a finite group with even order. Show that there exists an element with order 2 in $G$.
\end{problem}

\begin{problem}\label{problem-subgroup-of-cyclic-group-is-cyclic}
    Let $G$ be a cyclic group with generator $g$. Prove that any subgroup of $G$ must also be cyclic.\newline
    (\textit{Hint: Consider the solution for \myref{example-subgroups-of-Z}.})
\end{problem}

\begin{problem}\label{problem-subgroup-of-index-2}
    Let $G$ be a finite group. Suppose $H < G$ such that the index of $H$ in $G$ is 2. Prove that
    \begin{partquestions}{\roman*}
        \item $H \lhd G$;
        \item $H$ contains the squares of all elements of $G$; and
        \item an element $x \in G$ is in $H$ if $x$ has odd order.
    \end{partquestions}
\end{problem}

\newpage

\begin{problem}\label{problem-intersection-of-coprime-subgroups}
    Let $G$ be a finite group, $H \leq G$, and $K \leq G$. Suppose the greatest common divisor (GCD) of the order of $H$ and the order of $K$ is 1.
    \begin{partquestions}{\alph*}
        \item Show that the intersection of the groups $H$ and $K$ contains only the identity.
        \item Show that, if $H$ and $K$ are normal subgroups of $G$, then for any $h \in H$ and $k \in K$ we have $hk = kh$.
    \end{partquestions}
\end{problem}

\begin{problem}\label{problem-smallest-nonabelian-group}
    Let $G$ be a finite group with order $m$.
    \begin{partquestions}{\alph*}
        \item State the smallest value of $m$ such that $G$ is non-abelian.
        \item Prove that the value of $m$ found in \textbf{(a)} is the smallest value that allows $G$ to be non-abelian.
        \item Hence prove that for all even integers $n \geq m$, there exists a non-abelian group of order $n$.
    \end{partquestions}
\end{problem}

\begin{problem}\label{problem-quotient-of-group-mod-center-is-cyclic-implies-abelian}
    Let $G$ be a group, and suppose $G / \Z{G}$ is cyclic. Prove that $G$ is abelian.
\end{problem}
