\input{../SharedCommands.tex}
\usepackage{xr}

%=========== Global Variables ============
\newcommand{\version}{0.1}
\newcommand{\volumenumber}{2}
\newcommand{\volumename}{Rings}

% Quote page variables
\newcommand{\quotepagetext}{
    [Some] of the major discoveries in ring theory have helped shape the course of development of modern abstract algebra... A course in ring theory is an indispensable part of the education of any fledgling algebraist.
}
\newcommand{\quotepageattribution}{Tsit-Yuen Lam, 2001}
\newcommand{\quotepagecitation}{\cite{lam_2001}}

% Preface variables
\newcommand{\prefacevolumetext}{
    This volume covers the basics of ring theory. %TODO: Add
}

%============= Formatting ================
\linespread{1.05}

%============== Resources ================
\externaldocument{../Volume 0/AbstractAlgebraVol0}

%=========================================
\begin{document}
\frontmatterpages

%=========================================
\chapter{Introduction to Rings}
\section{Basic Algebraic Structures}
Before we introduce rings, we look at `simpler' algebraic structures and build our way up to rings.

\begin{definition}
    A \textbf{magma} is a set $M$ together with a binary operation $\ast$ which is closed. That is, if $a$ and $b$ are in $M$, then $a \ast b \in M$.
\end{definition}

\begin{example}
    Let $S$ be a set. Let $f: S \times S \to S$. Define the operation $\ast$ by $a \ast b = f(a, b)$. Then $(S, \ast)$ forms a magma since $a\ast b = f(a,b) \in S$.
\end{example}

\begin{definition}
    A \textbf{semigroup} $\mathcal{S}$ is a magma where the operation $\ast$ is associative. That is, $a\ast(b\ast c) = (a\ast b)\ast c$.
\end{definition}

\begin{example}
    Consider the magma $(S, \ast)$ given above. If the function $f$ above satisfies $f(a, f(b, c)) = f(f(a, b), c)$ then $(S, \ast)$ is a semigroup since
    \[
          a\ast(b\ast c) = f(a, f(b, c)) = f(f(a, b), c) = (a\ast b)\ast c.
    \]
\end{example}

\begin{definition}
    A \textbf{monoid} $M$ is a semigroup with an element $e$, called the \textbf{identity}, such that
    \[
        e \ast m = m \ast e = m
    \]
    for all $m \in M$.
\end{definition}

\begin{example}
    Let $X$ be a set. Let the set
    \[
        S = \{f \vert  f: X \to X\}.
    \]
    and let $\circ$ denote function composition. Then $(S, \circ)$ forms a monoid:
    \begin{itemize}
        \item \textbf{Closure}: If $f, g \in S$ then $f\circ g \in S$ since function composition is closed.
        \item \textbf{Associative}: Function composition is associative.
        \item \textbf{Identity}: The identity function $\id: X \to X, x\mapsto x$ is inside $S$ and
        \[
            \id \circ f = f \circ \id = f
        \]
        for all $f \in S$.
    \end{itemize}
    Thus $(S, \circ)$ forms a monoid.
\end{example}

\begin{definition}
    A \textbf{group} $G$ is a monoid where every element has an inverse. That is to say, for every $g \in G$, there exists $g^{-1} \in G$ such that
    \[
        g \ast g^{-1} = g^{-1} \ast g = e
    \]
    where $e$ is the identity in $G$.
\end{definition}
We have previously looked at many examples of groups, and have dedicated the previous part of this book looking at groups. We are now ready to look at \textbf{rings}, which is a type of algebraic structure with more conditions attached to it.

\newpage

\begin{definition}
    A \textbf{ring} is a set $R$ with two binary operations $\oplus$ and $\otimes$ satisfying the following axioms:
    \begin{itemize}
        \item \textbf{Addition-Abelian}: $(R, \oplus)$ forms an abelian group.
        \item \textbf{Multiplication-Monoid}: $(R, \otimes)$ forms a monoid.
        \item \textbf{Distributive}: $\otimes$ is distributive over $\oplus$. That is,
        \begin{itemize}
            \item $a \otimes (b \oplus c) = (a \otimes b) \oplus (b \otimes c)$; and
            \item $(a \oplus b) \otimes c = (a \otimes c) \oplus (b \otimes c)$.
        \end{itemize}
    \end{itemize}
    One may denote such a ring by $(R, \oplus, \otimes)$.
\end{definition}
\begin{remark}
    We often write $\oplus$ using the standard addition symbol $+$. The multiplication symbol $\otimes$ is usually omitted. This means that $x \otimes y$ is usually written as $xy$.
\end{remark}

%TODO ADD

\begin{exercise}
    Prove using the ring axioms that $\mathbb{Z}$ is a ring under addition and multiplication.\newline
    (\textit{You do \textbf{not} need to prove the \textbf{Distributive} axiom.})
\end{exercise}

For rings, we denote the additive identity using $0$ and the multiplicative identity using $1$.

%=========================================
\appendix
\chapter{Exercise Solutions}

%=========================================
\chapter{Problem Solutions}

%=========================================
\chapter{Image Acknowledgements}
Unless otherwise stated, all images are the author's own work, and are released under the same licence as this book.

%=========================================
\printbibliography[heading=bibintoc, title={References and Bibliography}]

\end{document}
