\section{Domains and Factorization}
\subsection*{Exercises}
\begin{questions}
    \item For the forward direction, suppose $a \vert b$ and $b \vert a$. Then there exist elements $u, v \in D$ such that $b = au$ and $a = bv$. So
    \[
        a = bv = (au)v = auv
    \]
    which means $uv = 1$ by cancellation law (\myref{prop-domain-cancellation-law}). Thus $u$ (and also $v$) is a unit, which shows that $a$ and $b$ are associates.

    For the reverse direction, if $a$ and $b$ are associates, then there exists a unit $u \in D$ such that $a = bu$, which also means that $b = au^{-1}$. Therefore one clearly sees that $a \vert b$ and $b \vert a$.

    \item Suppose $r$ is an irreducible and $q$ an associate of $r$, meaning there exists a unit $u$ in the integral domain (say $D$) such that $q = ru$. Suppose on the contrary that $q$ is reducible, meaning $q = ab$ where neither $a$ nor $b$ are units in $D$. Then $ru = ab$ which means $r = (u^{-1}a)b$. As $r$ is irreducible we know that either $b$ is a unit (contradicting the assumption) or $u^{-1}a$ is a unit. Since the product of units is a unit, thus $u(u^{-1}a) = a$ is a unit, contradicting the assumption. Therefore $q = ru$ is irreducible.

    \item \begin{partquestions}{\alph*}
        \item Note $2 = (1+i)(1-i)$ and $5 = (2+i)(2-i)$, and one can easily verify that each of these factors do not have a norm of $\pm1$. Therefore 2 and 5 are reducible in $\Z[i]$, so they cannot be prime (contrapositive of \myref{thrm-in-integral-domain-primes-are-irreducibles}).

        \item We first note $N(x) \geq 0$ for all $x \in \Z[i]$.
        \begin{partquestions}{\roman*}
            \item Without loss of generality we consider only $a$. We note that
            \begin{itemize}
                \item if $a \equiv 0 \pmod4$ then $a^2 \equiv 0 \pmod4$;
                \item if $a \equiv 1 \pmod4$ then $a^2 \equiv 1 \pmod4$;
                \item if $a \equiv 2 \pmod4$ then $a^2 \equiv 4 \equiv 0 \pmod4$; and
                \item if $a \equiv 3 \pmod4$ then $a^2 \equiv 9 \equiv 1 \pmod4$.
            \end{itemize}
            So the sum of two squares modulo 4 can only be 0, 1, or 2. Thus it is impossible for $a^2+b^2 \equiv 3\pmod4$.

            \item Seeking a contradiction, suppose $3 = pq$ where $p,q\in\Z[i]$ are both non-units. Then
            \[
                9 = N(3) = N(pq) = N(p)N(q).
            \]
            Since $p$ and $q$ are non-units we must have $N(p) = N(q) = 3$ (since otherwise $N(p) = 1$ which means $p$ is a unit by \myref{prop-properties-of-quadratic-integer-norm}). So if $p = a+bi$ then $a^2+b^2 = 3$. But we just proved in \textbf{(b)(i)} that this is impossible. Thus 3 is irreducible.

            \item Let $x = m+ni$, so $N(x) = m^2 + n^2$. Given $9 \vert N(x)$, we know $m^2 + n^2 \equiv 0 \pmod9$. We note the possibilities of squares modulo 9.
            \begin{table}[H]
                \centering
                \begin{tabular}{|l|l|l|}
                    \hline
                    $k\mod9$ & $k^2$ & $k^2\mod9$ \\ \hline
                    0 & 0 & 0 \\ \hline
                    1 & 1 & 1 \\ \hline
                    2 & 4 & 4 \\ \hline
                    3 & 9 & 0 \\ \hline
                    4 & 16 & 7 \\ \hline
                    5 & 25 & 7 \\ \hline
                    6 & 36 & 0 \\ \hline
                    7 & 49 & 4 \\ \hline
                    8 & 64 & 1 \\ \hline
                \end{tabular}
            \end{table}
            One may then verify that the set of pairs of $(m,n)$ that results in $m^2 + n^2 \equiv 0 \pmod9$ is
            \[
                \left\{(0,0),(0,3),(0,6),(3,0),(3,3),(3,6),(6,0),(6,3),(6,6)\right\}.
            \]
            In each case, both $m$ and $n$ are divisible by 3. Say $m = 3p$ and $n = 3q$ for some integers $p$ and $q$. Then $x = m + ni = 3p + (3q)i = 3(p+qi)$ which therefore means $3 \vert x$.

            \item Suppose $3 \vert uv$ for some $u,v\in D$. This means $uv = 3w$ for some $w \in D$. Note
            \[
                N(u)N(v) = N(uv) = N(3w) = 9N(w)
            \]
            and $9 = 3\times3$. We have 3 cases.
            \begin{enumerate}[label=\arabic*.]
                \item If $9 \vert N(u)$ we may use \textbf{(b)(iii)} to see that $3 \vert u$.
                \item Otherwise, if $9 \vert N(v)$ we may again use \textbf{(b)(iii)} to see that $3 \vert v$.
                \item Otherwise, we must have $3 \vert N(u)$ and $3 \vert N(v)$. Let's focus on $3 \vert N(u)$ only; let $u = x+yi$ for some integers $x$ and $y$. Then $3 \vert x^2+y^2$. Note for any integer $n$,
                \begin{itemize}
                    \item $n \equiv 0 \pmod3$ means $n^2 \equiv 0 \pmod3$;
                    \item $n \equiv 1 \pmod3$ means $n^2 \equiv 1 \pmod3$; and
                    \item $n \equiv 2 \pmod3$ means $n^2 \equiv 4 \equiv 1 \pmod3$.
                \end{itemize}
                So for $x^2+y^2$ to be divisible by 3, both $x$ and $y$ must be multiples of 3. Let $x = 3m$ and $y = 3n$ for some integers $n$; we see
                \[
                    N(u) = x^2+y^2 = (3m)^2 + (3n)^2 = 9(m^2+n^2)
                \]
                which means $9 \vert N(u)$. Finally use \textbf{(b)(iii)} to see that $3 \vert u$.
            \end{enumerate}
            Hence in any case, $3 \vert uv$ means $3 \vert u$ or $3 \vert v$, meaning 3 is prime.
        \end{partquestions}
    \end{partquestions}

    \item Suppose $N(x) = p$ where $p$ is a prime number, and $x = ab$ where $a,b \in \Z[\sqrt{d}]$. Taking norms yields $N(x) = N(ab) = N(a)N(b) = p$. But since $p$ is prime, one of $N(a)$ or $N(b)$ is 1, which by \myref{prop-properties-of-quadratic-integer-norm}, statement 3, we know that $a$ or $b$ is a unit. Therefore $x$ is irreducible.

    \item Suppose $\frac{a_1}{b_1} = \frac{a_2}{b_2}$ and $\frac{c_1}{d_1} = \frac{c_2}{d_2}$. We show that $\frac{a_1}{b_1} \times \frac{c_1}{d_1} = \frac{a_2}{b_2} \times \frac{c_2}{d_2}$, i.e. $\frac{a_1c_1}{b_1d_1} = \frac{a_2c_2}{b_2c_2}$, which is equivalent to showing that $(a_1c_1, b_1d_1) \mathrel{\sim} (a_2c_2, b_2d_2)$ by \myref{thrm-equivalence-class-equivalence}, i.e.
    \[
        a_1c_1b_2d_2 = b_1d_1a_2c_2.
    \]
    As $\frac{a_1}{b_1} = \frac{a_2}{b_2}$, we know that $(a_1, b_1) \mathrel{\sim} (a_2, b_2)$ (\myref{thrm-equivalence-class-equivalence}) which means $a_1b_2 = b_1a_2$. Similarly, because $\frac{c_1}{d_1} = \frac{c_2}{d_2}$ thus $c_1d_2 = d_1c_2$. Hence
    \begin{align*}
        a_1c_1b_2d_2 &= a_1b_2c_1d_2\\
        &= (a_1b_2)(c_1d_2)\\
        &= (b_1a_2)(d_1c_2)\\
        &= b_1d_1a_2c_2.
    \end{align*}
    so we have shown that $\frac{a_1}{b_1} \times \frac{c_1}{d_1} = \frac{a_2}{b_2} \times \frac{c_2}{d_2}$, meaning multiplication is well-defined.

    We now show that multiplication is commutative. Let $\frac ab, \frac cd \in \Frac{D}$. Then
    \begin{align*}
        \frac ab \times \frac cd &= \frac{ac}{bd}\\
        &= \frac{ca}{db} & (\text{since } D \text{ is commutative})\\
        &= \frac cd \times \frac ab
    \end{align*}
    so multiplication is commutative.

    \item \begin{partquestions}{\alph*}
        \item Disprove. One sees $1 \in \Z[x]$ and $x \in \Z[x]$ but $\frac1x \notin \Q[x]$.

        \item Prove. Define $\phi: \Frac{k\Z} \to \Q$ by $\phi([(ka, kb)]) = \frac ab$. We show that this is a well-defined isomorphism.
        \begin{itemize}
            \item \textbf{Well-defined}: Suppose $[(ka, kb)], [(kc, kd)] \in \Frac{k\Z}$ where $[(ka,kb)]=[(kc,kd)]$. Then $(ka, kb) \mathrel{\sim} (kc, kd)$ by \myref{thrm-equivalence-class-equivalence}, i.e. $k^2ad = k^2bc$. Hence one sees clearly that $\frac ab = \frac cd$. Therefore,
            \[
                \phi([(ka,kb)]) = \frac ab = \frac cd = \phi([(kc, kd)])
            \]
            and so $\phi$ is well-defined.

            \item \textbf{Homomorphism}: Let $[(ka, kb)], [(kc, kd)] \in \Frac{k\Z}$. Note
            \begin{align*}
                \phi([(ka,kb)] + [(kc,kd)]) &= \phi([(k^2(ad+bc), k^2bd)])\\
                &= \frac{k(ad+bc)}{kbd}\\
                &= \frac{k^2(ad+bc)}{k^2bd}\\
                &= \frac {ka}{kb} + \frac {kc}{kd}\\
                &= \phi([(ka,kb)]) + \phi([(kc,kd)])
            \end{align*}
            and
            \begin{align*}
                \phi([(ka,kb)] \times [(kc,kd)]) &= \phi([(k^2ac, k^2bd)])\\
                &= \frac{kac}{kbd}\\
                &= \frac{k^2ac}{k^2bd}\\
                &= \frac {ka}{kb} \times \frac {kc}{kd}\\
                &= \phi([(ka,kb)]) \times \phi([(kc,kd)])
            \end{align*}
            which shows that $\phi$ is a homomorphism.

            \item \textbf{Injective}: Let $[(ka, kb)], [(kc, kd)] \in \Frac{k\Z}$ such that $\phi([(ka, kb)]) = \phi([(kc, kd)])$. Then $\frac ab = \frac cd$, meaning $ad = bc$. Hence $(ka, kb) \mathrel{\sim} (kc, kd)$ by definition of the equivalence relation on $\Frac{k\Z}$, and so $[(ka, kb)] = [(kc, kd)]$ by \myref{thrm-equivalence-class-equivalence} again. Hence $\phi$ is injective.

            \item \textbf{Surjective}: Suppose $\frac ab \in \Q$ where $a, b \in \Z$ such that $b \neq 0$. Then note that $[(ka, kb)]$ is its pre-image since
            \[
                \phi([(ka, kb)]) = \frac ab
            \]
            which proves that $\phi$ is surjective.
        \end{itemize}
        Therefore $\phi$ is a well-defined ring isomorphism, meaning $\Frac{k\Z} \cong \Q$.
    \end{partquestions}

    \item For the forward direction, we prove by contrapositive. If $a$ is a unit then $\princ{b} = \princ{ab}$ by \myref{prop-principal-ideals-equal-iff-associates}, so $\princ{ab} \not\subset \princ{b}$.

    For the reverse direction, we again consider contrapositive. Assume $\princ{ab}$ is not a proper subset of $\princ{b}$, meaning $\princ{b} \subseteq \princ{ab}$. But one clearly sees that $\princ{ab} \subseteq \princ{b}$. Thus $\princ{b} = \princ{ab}$ which means $a$ is a unit, again by \myref{prop-principal-ideals-equal-iff-associates}.

    \item We prove the contrapositive of the forward direction. Suppose $f(x)$ is reducible in $D[x]$. Then, since $f(x)$ is primitive, $f(x) = p(x)q(x)$ for some non-unit polynomials $p(x), q(x) \in D[x]$. Therefore $p(x), q(x) \in \Frac{D}$, which are non-units in $\Frac{D}$. Thus $f(x) = p(x)q(x)$ for some non-unit polynomials in $p(x), q(x) \in \Frac{D}$, so $f(x)$ is reducible in $\Frac{D}$.

    We now prove the contrapositive of the reverse direction. Suppose $f(x)$ is reducible in $\Frac{D}$, so there exists non-constant polynomials $p(x), q(x) \in \Frac{D}$ such that $f(x) = p(x)q(x)$. Then there exists polynomials $P(x), Q(x) \in D[x]$ such that $f(x) = P(x)Q(x)$ and $\deg P(x) = \deg p(x)$ and $\deg Q(x) = \deg q(x)$ by \myref{lemma-reducible-in-field-of-fractions-means-reducible-in-UFD}. Hence $f(x)$ is reducible in $D[x]$.

    \item For a field $F$, choose the norm to be $N(x) = 1$ for any $x \in F$. Clearly \textbf{EF1} is satisfied since $1 = N(x) \leq N(xy) = 1$ for any $x,y\in F$. For \textbf{EF2}, for any $n, d \in F$, we can write $n = (nd^{-1})d + 0$. Note $nd^{-1}$ exists since we are in a field, so \textbf{EF2} is satisfied. Therefore $F$ is a Euclidean domain.

    \item We note
    \begin{itemize}
        \item all fields are Euclidean domains by \myref{thrm-field-is-euclidean-domain};
        \item all Euclidean domains are PIDs by \myref{thrm-euclidean-domain-is-PID}; and
        \item all PIDs are UFDs by \myref{thrm-PID-is-UFD}.
    \end{itemize}
\end{questions}

\subsection*{Problems}
\begin{questions}
    \item Given that we have integers $a$ and $b$ such that $a^2 + b^2$ is a prime, therefore $N(a+bi) = a^2+b^2$ is prime. Therefore $a+bi$ is irreducible (\myref{prop-properties-of-quadratic-integer-norm}). Since $\Z[i]$ is a Euclidean domain (\myref{example-gaussian-integers-is-euclidean-domain}) it is thus a PID (\myref{thrm-euclidean-domain-is-PID}) and so any irreducible in $\Z[i]$ is prime (\myref{thrm-in-PID-prime-iff-irreducible}).

    \item We induct on $n$. When $n = 2$, this reduces down to the definition of a prime element in an integral domain. Now suppose $p \vert a_1a_2\cdots a_k$ for some $k \geq 2$. We are to show that this works for $k + 1$.

    Note $p \vert a_1a_2\cdots a_ka_{k+1}$ is the same as $p \vert (a_1a_2\cdots a_k)a_{k+1}$. By definition of a prime element we know that $p \vert a_1a_2\cdots a_k$ or $p \vert a_{k+1}$. Therefore, using induction hypothesis on the former, we see $p\vert a_1$, or $p\vert a_2$, or $p \vert a_3$, etc., or $p \vert a_k$, or $p \vert a_{k+1}$, which proves the statement for $k + 1$.

    Therefore by induction we establish the required result.

    \item Since $\Z[i]$ is a PID (as shown earlier), any irreducible in $\Z[i]$ is prime (\myref{thrm-in-PID-prime-iff-irreducible}). So we just need to show whether the following are reducible or irreducible.
    \begin{partquestions}{\alph*}
        \item $6 = 2 \times 3$. As $N(2) = 4 \neq 1$ and $N(3) = 9 \neq 1$, neither 2 nor 3 are units in $\Z[i]$. Therefore 6 is reducible, and so it is not prime.

        \item We claim 7 is irreducible. Suppose $7 = wz$ for some $w,z \in \Z[i]$. Then $49 = N(wz) = N(w)N(z)$. Suppose neither $w$ nor $z$ are units, which means $N(w) = N(z) = 7$.

        If $w = a+bi$ for some integers $a$ and $b$, then $N(w) = a^2+b^2 = 7$. This means $a^2 = 7 - b^2$. For a solution to exist the right hand side cannot be negative, so we note that $b$ can only be -2, -1, 0, 1, or 2. For each one of these possibilities we see that $7 - b^2$ is not a square, so there is no solution for $w$ (and likewise for $z$).

        Hence one of $N(w)$ or $N(z)$ is 1, meaning that one of $w$ or $z$ is a unit. Hence 7 is irreducible and therefore prime.

        \item 13 is not since $13 = (3+2i)(3-2i)$, and one sees that $N(3+2i) = N(3-2i) = 13 \neq 1$, so neither $3+2i$ nor $3-2i$ are units. Hence 13 is reducible and therefore not prime.

        \item $1+2i$ is irreducible since $N(1+2i) = 5$ is prime (\myref{prop-properties-of-quadratic-integer-norm}). Hence $1+2i$ is prime.

        \item $3+2i$ is reducible since $3+4i = (2+i)^2$ and $N(2+i) = 5 \neq 1$, so $2+i$ is not a unit. Hence $3+2i$ is not prime.

        \item $5+i$ is reducible since $5+i = (1+i)(3-2i)$ and $N(1+i) = 2 \neq 1$ and $N(3-2i) = 13 \neq 1$, so neither $1+i$ nor $3-2i$ are units. Hence $5+i$ is not a prime.
    \end{partquestions}

    \item To prove that $R$ is an equivalence relation, we need to show that $R$ is reflexive, symmetric, and transitive.
    \begin{itemize}
        \item \textbf{Reflexive}: Clearly for any $a \in D$ we see $a = 1a$. 1 is a unit since it is its own inverse. Hence $a$ is its own associate, so $a\mathrel{R}a$.

        \item \textbf{Symmetric}: Suppose $a,b \in D$ such that $a\mathrel{R}b$. Then $a = ub$ for some unit $u \in D$. Then $b = u^{-1}a$, and since $u^{-1}$ is a unit, thus $b$ and $a$ are associates. Hence $b\mathrel{R}a$.

        \item \textbf{Transitive}: Suppose $a,b,c\in D$ such that $a\mathrel{R}b$ and $b\mathrel{R}c$. Then there exist units $u,v \in D$ such that $a = ub$ and $b = vc$. Thus $a = uvc$. As the product of units is a unit \myref{prop-product-of-units-is-unit}, therefore $uv$ is a unit and so $a$ and $c$ are associates. Hence $a\mathrel{R}c$.
    \end{itemize}
    Therefore $R$ is an equivalence relation.

    \item We see
    \begin{align*}
        (3x+2)(x+4) &= 3x^2 + 12x + 2x + 8\\
        &= 3x^2 + 14x + 8\\
        &= 3x^2 + 4x + 3
    \end{align*}
    in $\Z_5[x]$, and we also see
    \begin{align*}
        (4x+1)(2x+3) &= 8x^2 + 12x + 2x + 3\\
        &= 8x^2 + 14x + 3\\
        &= 3x^2 + 4x + 3
    \end{align*}
    in $\Z_5[x]$, which is the same polynomial as before.

    We note that this does not contradict \myref{corollary-polynomial-ring-over-field-is-UFD} since the factors, up to reordering, are associates. Note
    \begin{align*}
        3x+2 = 8x+2 = 2(4x+1)\\
        x+4 = 6x+9 = 3(2x3)
    \end{align*}
    and since $\Z_5$ is a field, so 2 and 3 are units as all non-zero elements in $\Z_5$ are units. Therefore $3x+2$ and $4x+1$ are associates, as are $x+4$ adn $2x+3$.

    \item For the forward direction, suppose $p$ is a prime. Let $a, b \in D$ such that $ab \in \princ{p}$, which means there exists a $k \in D$ such that $ab = kp$. So $p \vert ab$. As $p$ is a prime, this means that either $p \vert a$ or $p \vert b$. Without loss of generality assume $p \vert a$, meaning $a = qp$ for some $q \in D$. Hence $a \in \princ{p}$, which shows that $\princ{p}$ is a prime ideal.

    For the reverse direction, suppose $\princ{p}$ is prime and $p \vert ab$. Thus $ab = kp$ for some $k \in D$, which means $ab \in \princ{p}$. So $a \in \princ{p}$ or $b \in \princ{p}$ by definition of prime ideal. Without loss of generality assume $a \in \princ{p}$; we see $a = qp$ for some $q \in D$, meaning $p \vert a$. Hence $p$ is a prime element.

    \item Define $\phi: \Frac{F} \to F$ such that $[(a,b)]\mapsto ab^{-1}$. We show that $\phi$ is a well-defined ring isomorphism.
    \begin{itemize}
        \item \textbf{Well-defined}: Let $[(a,b)], [(c,d)] \in \Frac{F}$ and suppose $[(a,b)] = [(c,d)]$. Then $(a,b) \mathrel{\sim} (c,d)$ by \myref{thrm-equivalence-class-equivalence}, which means $ad = bc$ by definition of the equivalence relation on the field of fractions. Hence $ab^{-1} = cd^{-1}$ (remembering that the operation is commutative in a field), so
        \[
            \phi([(a,b)]) = ab^{-1} = cd^{-1} = \phi([(c,d)]),
        \]
        meaning that $\phi$ is well-defined.

        \item \textbf{Homomorphism}: Let $[(a,b)], [(c,d)] \in \Frac{F}$. Note
        \begin{align*}
            \phi([(a,b)] + [(c,d)]) &= \phi([(ad+bc, bd)])\\
            &= (ad+bc)(bd)^{-1}\\
            &= (ad)(bd)^{-1} + (bc)(bd)^{-1}\\
            &= add^{-1}b^{-1} + bcd^{-1}b^{-1}\\
            &= ab^{-1} + cd^{-1}\\
            &= \phi([(a,b)]) + \phi([(c,d)])
        \end{align*}
        and
        \begin{align*}
            \phi([(a,b)][(c,d)]) &= \phi([(ac,bd)])\\
            &= (ac)(bd)^{-1}\\
            &= (ab^{-1})(cd^{-1})\\
            &= \phi([(a,b)])\phi([(c,d)])
        \end{align*}
        so $\phi$ is a ring homomorphism.

        \item \textbf{Injective}: Let $[(a,b)], [(c,d)] \in \Frac{F}$ such that $\phi([(a,b)]) = \phi([(c,d)])$. Then $ab^{-1} = cd^{-1}$, so $ad = bc$. Hence $(a,b) \mathrel{\sim} (c,d)$ by definition of the equivalence relation on the field of fractions, so $[(a,b)] = [(c,d)]$ by \myref{thrm-equivalence-class-equivalence}.

        \item \textbf{Surjective}: Suppose $r \in F$. Note $r1^{-1} = r$. Therefore $\phi([(r, 1)]) = r1^{-1} = r$, meaning that every $r \in F$ has a pre-image under $\phi$.
    \end{itemize}
    Therefore $\phi$ is a well-defined ring isomorphism, meaning that $\Frac{F} \cong F$.

    \item We note that the only units of $\Z[\sqrt{-6}]$ are $\pm1$. We claim that 2, 3, and $\sqrt{-6}$ are all irreducible in $\sqrt{-6}$.

    We first show that 2 is irreducible in $\Z[\sqrt{-6}]$. Suppose $2 = (a+b\sqrt{-6})(c+d\sqrt{-6})$ for some integers $a,b,c,d\in \Z$. So
    \[
        4 = N(2) = N((a+b\sqrt{-6})(c+d\sqrt{-6})) = (a^2+6b^2)(c+6d^2)
    \]
    which means that $a^2+6b^2$ must be 1, 2, or 4.
    \begin{itemize}
        \item If $a^2+6b^2 = 1$, then one sees that we must have $a = \pm1$ and $b = 0$, which means $a+b\sqrt{-6} = \pm1$ is a unit.
        \item If $a^2+6b^2 = 2$ then $a^2 = 2 - 6b^2$. As $b^2 \geq 0$, we see that $b = 0$ is the only way for the right hand side to not be negative. But $a^2 = 2$ has no solution. Thus $a^2+6b^2 = 2$ is impossible.
        \item If $a^2+6b^2 = 4$ then $c^2+6d^2 = 1$. We may then use the argument for the first case to show that $c+d\sqrt{-6} = \pm1$ which is a unit.
    \end{itemize}
    Therefore 2 is an irreducible.

    We next show that 3 is irreducible in $\Z[\sqrt{-6}]$. Suppose $3 = (a+b\sqrt{-6})(c+d\sqrt{-6})$ for some integers $a,b,c,d\in \Z$. So
    \[
        9 = N(3) = N((a+b\sqrt{-6})(c+d\sqrt{-6})) = (a^2+6b^2)(c+6d^2)
    \]
    which means that $a^2+6b^2$ must be 1, 3, or 9.
    \begin{itemize}
        \item If $a^2+6b^2 = 1$, then $a+b\sqrt{-6} = \pm1$ is a unit by previous working.
        \item If $a^2+6b^2 = 3$ then $a^2 = 3 - 6b^2$. As $b^2 \geq 0$, we see that $b = 0$ is the only way for the right hand side to not be negative. But $a^2 = 3$ has no solution. Thus $a^2+6b^2 = 3$ is impossible.
        \item If $a^2+6b^2 = 4$ then $c^2+6d^2 = 1$ which means $c+d\sqrt{-6} = \pm1$ which is a unit.
    \end{itemize}
    Therefore 3 is an irreducible.

    We finally show that $\sqrt{-6}$ is an irreducible in $\Z[\sqrt{-6}]$. Suppose $\sqrt{-6} = (a+b\sqrt{-6})(c+d\sqrt{-6})$ for some integers $a,b,c,d\in \Z$. So
    \[
        6 = N(\sqrt{-6}) = N((a+b\sqrt{-6})(c+d\sqrt{-6})) = (a^2+6b^2)(c+6d^2)
    \]
    which means that $a^2+6b^2$ must be 1, 2, 3, or 6.
    \begin{itemize}
        \item If $a^2+6b^2 = 1$, then $a+b\sqrt{-6} = \pm1$ is a unit.
        \item If $a^2+6b^2 = 2$ is impossible by earlier working.
        \item If $a^2+6b^2 = 3$ is impossible by earlier working.
        \item If $a^2+6b^2 = 6$ then $c^2+6d^2 = 1$ so $c+d\sqrt{-6} = \pm1$ which is a unit.
    \end{itemize}
    Therefore $\sqrt{-6}$ is an irreducible.

    Clearly 2, 3, and $\sqrt{-6}$ are not associates of each other.

    Finally, notice $6 = 2 \times 3 = (-\sqrt{-6})(\sqrt{-6})$ which is two factorizations of 6 into irreducibles. Therefore $\Z[\sqrt{-6}]$ is not a UFD.

    \item For the forward direction, assume $p$ is an irreducible, and let $p \vert ab$ for some $a,b \in D$. Therefore $ab = pq$ for some $q \in D$. As $D$ is a UFD, we may factorise $a$, $b$, and $q$ into irreducibles, say $a_1, a_2, \dots, a_r$, $b_1, b_2, \dots, b_s$, and $q_1, q_2, \dots, q_t$ respectively. Therefore
    \[
        a_1a_2\cdots a_r b_1b_2\cdots b_s = p q_1q_2\cdots q_t.
    \]
    As $p$ is also an irreducible, $p$ must be a $a_i$ or $b_i$, which means $p$ divides $a$ or $b$.

    For the reverse direction, since UFDs are integral domains, therefore any prime is an irreducible (\myref{thrm-in-integral-domain-primes-are-irreducibles}).

    \item \begin{partquestions}{\alph*}
        \item Suppose $a$ and $b$ are associates in $D$. So there is a unit $u \in D$ such that $a = ub$. Note
        \begin{align*}
            N(a) &= N(ub) \leq N(b) \text{ and}\\
            N(b) &= N(u^{-1}a) \leq N(a)
        \end{align*}
        by \textbf{EF1}. Hence $N(a) = N(b)$.

        \item For the forward direction, suppose $u \in D$ is a unit. Then there exists a $v \in D$ such that $uv = 1$. Therefore $N(uv) = N(1)$. Note that
        \begin{align*}
            N(u) &\leq N(uv) = N(1) \text{ and}\\
            N(1) &\leq N(u1) = N(u)
        \end{align*}
        by \textbf{EF1}, so $N(u) = N(1)$.

        For the reverse direction, suppose $N(u) = N(1)$. By \textbf{EF2}, there exists $q, r \in D$ such that $1 = uq + r$ with $r = 0$ or $N(r) < N(u)$. As $N(u) = N(1)$, so $r = 0$ or $N(r) < N(1)$. But $N(1) \leq N(1x) = N(x)$ for all $x \in D$, so $N(1) > N(r)$ is impossible. Hence $r = 0$, meaning $uq = 1$, which shows that $u$ is a unit.
    \end{partquestions}

    \item We first show that $\phi(1) = 1$. Let $\phi(1) = a$ for some $a \in \mathbb{N} \cup \{0\}$. Since $1 = 1 \times 1$, therefore
    \[
        a = \phi(1) = \phi(1 \times 1) = \phi(1)\phi(1) = a^2
    \]
    which shows $a^2 = a$. As we are in $\mathbb{N} \cup \{0\}$, thus $a = 0$ or $a = 1$. But if $\phi(1) = 0$ then we see
    \[
        \phi(x) = \phi(1x) = \phi(1)\phi(x) = 0\phi(x) = 0
    \]
    but $\phi$ is non-constant, a contradiction. Hence $\phi(1) = 1$.

    Now let $u$ be a unit in $D$, meaning that there exists a $v \in D$ such that $uv = 1$. So one sees
    \[
        1 = \phi(1) = \phi(uv) = \phi(u)\phi(v).
    \]
    As we working with the non-negative integers, the only way to a product to equal 1 we must have $\phi(u) = \phi(v) = 1$.

    \item Suppose there are only finitely many irreducible polynomials within $F[x]$, say $p_1(x), p_2(x), \dots, p_n(x)$. As $F[x]$ is a PID (\myref{thrm-polynomial-ring-over-field-is-a-PID}) thus these polynomials are also prime (\myref{thrm-in-PID-prime-iff-irreducible}).

    Consider the polynomial
    \[
        q(x) = (p_1(x) + 1)(p_2(x) + 1)\cdots(p_n(x) + 1).
    \]
    By construction we see that none of $p_1(x), p_2(x), \dots, p_n(x)$ divide $q(x)$. Therefore, either
    \begin{itemize}
        \item $q(x)$ is prime and thus irreducible (\myref{thrm-in-integral-domain-primes-are-irreducibles}), meaning that the list of irreducible polynomials is not complete; or
        \item there is another prime (and thus irreducible) polynomial that divides $q(x)$, which also means that the above list of irreducible polynomials is not complete.
    \end{itemize}
    Either way, this contradicts that there are a finite number of prime (and this irreducible) polynomials, meaning that there are infinitely many irreducible polynomials in $F[x]$.

    \item Let $I$ be a non-trivial ideal in $\Z[i]$. As $\Z[i]$ is a PID (it is a Euclidean domain which means it is a PID), suppose $I = \princ{a+bi}$ for some $a,b \in \Z$. Note that
    \[
        a^2 + b^2 = (a+bi)(a-bi) \in I.
    \]

    Now let $x+yi \in \Z[i]$. Euclid's division lemma (\myref{lemma-euclid-division}) tells us that there exist $q_1, q_2, r_1, r_2 \in \Z$ such that
    \begin{align*}
        x &= q_1(a^2+b^2) + r_1\\
        y &= q_2(a^2+b^2) + r_2
    \end{align*}
    where $0 \leq r_1, r_2 \leq a^2+b^2$. So
    \begin{align*}
        (x+yi)+I &= \left(\left(q_1(a^2+b^2) + r_1\right) + \left(q_2(a^2+b^2) + r_2\right)i\right) + I\\
        &= \left((a^2+b^2)(q_1+q_2i) + r_1 + r_2i\right) + I\\
        &= (r_1 + r_2i) + I & (\text{as } a^2+b^2 \in I)
    \end{align*}
    Notice that $(x+yi)+I = (r_1+r_2i)+I$ is an element of $\Z[i]/I$ and that there are only finitely many integers within the interval $[0, a^2+b^2]$. Therefore $\Z[i]/I$ is finite, since $0 \leq r_1, r_2 \leq a^2+b^2$.

    \item As $\Z[x]$ is a UFD (\myref{corollary-Z-is-UFD}) we note that the factorizations mentioned below are unique.
    \begin{partquestions}{\roman*}
        \item We note that if one die shows $n$, then the other die must have a value of $k - n$ in order for both dice's sum to be $k$. Similarly, in order to obtain a term with degree $k$, and if the first term to multiply is $x^n$, then the other has to be $x^{k-n}$ in order for their product to have degree $k$.

        \item $P(x) = x^2(x+1)^2(x^2+x+1)^2(x^2-x+1)^2$.

        We know that both $x$ and $x+1$ are irreducible by \myref{problem-primitive-degree-1-polynomial-in-Z[x]-is-irreducible}.

        $x^2+x+1$ is irreducible since it is equal to $(x+\frac12)^2 + \frac34$ which is never zero for all integers $x$, which means that $x^2+x+1$ has no zeroes in $\Z$ and so it is irreducible (\myref{thrm-degree-2-or-3-irreducible-iff-has-no-zeroes}).

        $x^2-x+1$ is irreducible since it is equal to $(x-\frac12)^2 + \frac34$ which is never zero for all integers $x$, which means that $x^2-x+1$ has no zeroes in $\Z$ and so it is irreducible (\myref{thrm-degree-2-or-3-irreducible-iff-has-no-zeroes}).

        Hence the factorization given here is indeed the unique factorization into irreducibles.

        \item Since the factorization in \textbf{(ii)} is into irreducibles, $f(x)$ must consist of the same irreducibles. Thus $f(x) = x^q(x+1)^r(x^2+x+1)^s(x^2-x+1)^t$.

        Note that $0 \leq q,r,s,t \leq 2$, since otherwise it would exceed the number of factors of the appropriate irreducible in the factorization of $P(x)$ obtained in \textbf{(ii)}.

        \item We see
        \[
            f(1) = 1^{a_1} + 1^{a_2} + \cdots + 1^{a_6} = 6
        \]
        and
        \[
            f(1) = 1^q(1+1)^r(1+1+1)^s(1-1+1)^t = 2^r3^s
        \]
        so $6 = 2^r3^s$. But as $6 = 2 \times 3$, therefore $r = s = 1$.

        We also note that
        \[
            f(0) = 0^{a_1} + 0^{a_2} + \cdots + 0^{a_6} = 0
        \]
        and
        \[
            f(0) = 0^q(0+1)^r(0+0+1)^s(0+0+1)^t = 0^q
        \]
        so $0^q = 0$. As we defined $0^0 = 1$, we see $q \neq 0$.

        \item So far we have $f(x) = x^q(x+1)(x^2+x+1)(x^2-x+1)^t$. If $q = 2$ then $f(x) = x^2(x+1)(x^2+x+1)(x^2-x+1)^t$. But since $P(x) = f(x)g(x)$ and the smallest sum of two dice is 2, that means one of the $b_i$ in $g(x)$ must be zero. But in \textbf{(iii)} we said that every $b_i$ is a positive integer, a contradiction. Hence $q \neq 2$.

        \item We deduce from \textbf{(iv)} and \textbf{(v)} that $q = 1$, meaning $f(x) = x(x+1)(x^2+x+1)(x^2-x+1)^t$. We now list all possibilities for $f(x)$.
        \begin{itemize}
            \item If $t = 0$ then $f(x) = x + 2x^2 + 2x^3 + x^4$, so one choice is 1, 2, 2, 3, 3, and 4.
            \item If $t = 1$ then $f(x) = x + x^2 + x^3 + x^4 + x^5 + x^6$, so one choice is 1, 2, 3, 4, 5, and 6 (which is the normal choice for a dice).
            \item If $t = 2$ then $f(x) = x + x^3 + x^4 + x^5 + x^6 + x^8$, so another choice is 1, 3, 4, 5, 6, and 8.
        \end{itemize}

        \item The labels of the other pair of dice are
        \[
            1,2,2,3,3,4 \quad\text{and}\quad 1,3,4,5,6,8
        \]
        which are known as Sicherman dice.
    \end{partquestions}
\end{questions}
